---
title: "3v2_TotNumSpec_plus1_ratioFC_Pval_compareTempsSameDay"
author: "Shelly Trigg"
date: "3/23/2019"
output: html_document
---


Load packages
```{r}
library(vegan)
library(ggplot2)
library(dplyr)
library(gtools)
```

Load Abacus data, parse out ADJNSAF values, and simplify column names to just sample number
```{r}
#upload data file
ABACUSdata <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/ABACUS_output021417.tsv", sep = "\t", header=TRUE, stringsAsFactors = FALSE)
#select only columns containing NUMSPECTOT and ADJNSAF and Protein ID
ABACUSdata_TOTNUMSPEC <- ABACUSdata[,c(1,grep("NUMSPECSTOT", colnames(ABACUSdata)))]
ABACUSdata_TOTNUMSPEC <- ABACUSdata_TOTNUMSPEC[,-grep("ALL_", colnames(ABACUSdata_TOTNUMSPEC))]

ABACUSdata_NSAF <- ABACUSdata[,c(1,grep("ADJNSAF", colnames(ABACUSdata)))]


## change column names in ABACUSdata to just sampleID
colnames(ABACUSdata_TOTNUMSPEC) <- gsub(pattern = "X20161205_SAMPLE_", "", colnames(ABACUSdata_TOTNUMSPEC))
colnames(ABACUSdata_TOTNUMSPEC) <- gsub(pattern = "_NUMSPECSTOT", "", colnames(ABACUSdata_TOTNUMSPEC))

colnames(ABACUSdata_NSAF) <- gsub(pattern = "X20161205_SAMPLE_", "", colnames(ABACUSdata_NSAF))
colnames(ABACUSdata_NSAF) <- gsub(pattern = "_ADJNSAF", "", colnames(ABACUSdata_NSAF))

```

Load meta data file with temperature and day information
```{r}
#upload meta data; this was a csv file I create from Rhonda's notebook entry: https://github.com/Ellior2/Ellior2.github.io/blob/master/_posts/2017-3-11-NMDS-analysis.md
meta_data <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/Rhonda_new_sample_names.csv", header = TRUE, stringsAsFactors = FALSE)
meta_data$silo <- substr(meta_data$Contents,5,5)
meta_data$day <- substr(meta_data$SampleName,5,6)
meta_data$SampleName <- gsub(pattern = "H","",meta_data$SampleName)
meta_data$SampleName <- gsub(pattern = "C","",meta_data$SampleName)
#create a temperature column
meta_data$temp <- "temp"
for(i in 1:nrow(meta_data)){
  if(meta_data$silo[i] == "2"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "3"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "9"){
    meta_data$temp[i] <- "29"
  }
  if(meta_data$silo[i] == "e"){
    meta_data$temp[i] <- "16"
  }
}
```

Reformat Abacus data for NMDS
```{r}
#Transpose- switch rows and columns
tABACUSdata_TOTNUMSPEC <- t.data.frame(ABACUSdata_TOTNUMSPEC[,-1])
colnames(tABACUSdata_TOTNUMSPEC) <- ABACUSdata_TOTNUMSPEC[,1]
tABACUSdata_TOTNUMSPEC <- cbind(data.frame(rownames(tABACUSdata_TOTNUMSPEC)),tABACUSdata_TOTNUMSPEC)
colnames(tABACUSdata_TOTNUMSPEC)[1] <- "SampleID"

#add meta data to abacus data
tABACUSdata_TOTNUMSPEC <- merge(meta_data[,c(1,2,7,8)],tABACUSdata_TOTNUMSPEC, by = "SampleID")

#Transpose- switch rows and columns
tABACUSdata_NSAF <- t.data.frame(ABACUSdata_NSAF[,-1])
colnames(tABACUSdata_NSAF) <- ABACUSdata_NSAF[,1]
tABACUSdata_NSAF <- cbind(data.frame(rownames(tABACUSdata_NSAF)),tABACUSdata_NSAF)
colnames(tABACUSdata_NSAF)[1] <- "SampleID"

#add meta data to abacus data
tABACUSdata_NSAF <- merge(meta_data[,c(1,2,7,8)],tABACUSdata_NSAF, by = "SampleID")



#add meta data to abacus data
#Remove Silo 2 and day 15
silo2_3_9 <- tABACUSdata_TOTNUMSPEC[which(tABACUSdata_TOTNUMSPEC$day != "15"),]
#make rownames from Sample ID column so that the NMDS knows what's what
rownames(silo2_3_9) <- silo2_3_9$SampleID
#order the data frame by day and temperature so coloring the points on the plot is easier
silo2_3_9 <- silo2_3_9[order(as.numeric(silo2_3_9$day),silo2_3_9$temp),]


silo2_3_9_NSAF <- tABACUSdata_NSAF[which(tABACUSdata_NSAF$day != "15"),]
#make rownames from Sample ID column so that the NMDS knows what's what
rownames(silo2_3_9_NSAF) <- silo2_3_9_NSAF$SampleID
#order the data frame by day and temperature so coloring the points on the plot is easier
silo2_3_9_NSAF <- silo2_3_9_NSAF[order(as.numeric(silo2_3_9_NSAF$day),silo2_3_9_NSAF$temp),]

```

Determine if any proteins have zero NUMSPECSTOT vals for all samples; or any that have zero ADJNSAF vals in all samples
```{r}
#totnum spec first
no_val_proteins <- silo2_3_9[,which(apply(silo2_3_9, 2, var) == 0)]
ncol(no_val_proteins)

#nsaf next
no_val_proteins_NSAF <- silo2_3_9_NSAF[,which(apply(silo2_3_9_NSAF, 2, var) == 0)]
ncol(no_val_proteins_NSAF)

```

Remove proteins with zero TOTNUMSPEC values in all samples
```{r}
silo2_3_9_nozerovar <- silo2_3_9[,-c(1:4,which(colnames(silo2_3_9) %in% colnames(no_val_proteins)))]
#check to make sure it worked
ncol(silo2_3_9)
ncol(silo2_3_9[,-c(1:4,which(colnames(silo2_3_9) %in% colnames(no_val_proteins)))])
ncol(silo2_3_9)-ncol(silo2_3_9_nozerovar)
```

Remove proteins with zero NSAF values in all samples
```{r}
silo2_3_9_nozerovar <- silo2_3_9_nozerovar[,which(!(colnames(silo2_3_9_nozerovar) %in% colnames(no_val_proteins_NSAF)))]
#check to make sure it worked
ncol(silo2_3_9_nozerovar[,which(!(colnames(silo2_3_9_nozerovar) %in% colnames(no_val_proteins_NSAF)))])
```

Sum replicate sample NUMSPECSTOT for each protein; this also adds 1 to all sums so no zero values will remain after (see end of code)
```{r}
df_sum <- data.frame()
#loop through the data and calculate the sum between replicates for each protein
for (i in seq(1,nrow(silo2_3_9_nozerovar),2)){
  #this calculates the sum for each odd number row and the row following it
  df_sum_row <- apply(silo2_3_9_nozerovar[c(i,i+1),],2,function(x) sum(x)+1)
  #this sequencially combines rows of data together after the sum is generated
  df_sum <- rbind(df_sum, df_sum_row)
}
#add column names to sum'd data
colnames(df_sum) <- colnames(silo2_3_9_nozerovar)
df_sum$SampleID <- rownames(silo2_3_9_nozerovar[-grep("A",rownames(silo2_3_9_nozerovar)),])
new_data_sum <- merge(meta_data[,c("SampleID", "silo", "day", "temp")],df_sum, by = "SampleID")
#order dataframe by day (earliest) and temp
new_data_sum<- new_data_sum[order(as.numeric(new_data_sum$day),new_data_sum$temp),]
```


Calculate sum of NUMSPECTOT for each sample
```{r}
spec_sum <- data.frame(apply(new_data_sum[,-c(1:4)], 1, sum))
spec_sum <- cbind(new_data_sum$SampleID,spec_sum)
colnames(spec_sum) <- c("SampleID", "sum")

#combine meta data with sum data
spec_sum <- merge(new_data_sum[,1:4], spec_sum, by = "SampleID")
#order dataframe by day (earliest) and temp
spec_sum <- spec_sum[order(as.numeric(spec_sum$day), spec_sum$temp),]

```


Calculate ratio of Protein NUMSPECSTOT / Sample NUMSPECSTOT
```{r}
df_ratio <- data.frame()
#loop through the data and calculate the sum between replicates for each protein
for (i in 1:nrow(spec_sum)){
  #this calculates the sum for each odd number row and the row following it
  for(j in 1:ncol(new_data_sum[,-c(1:4)])){
    df_ratio[i,j] <- new_data_sum[i,j+4]/spec_sum[i,"sum"]*100000
  }  
}
#add column names to ratio data
colnames(df_ratio) <- colnames(new_data_sum[,-c(1:4)])
df_ratio <- cbind(new_data_sum[,1:4], df_ratio)
```


See what the lowest values are
```{r}
head(unique(sort(unlist(df_ratio[,-c(1:4)]))))
```

order by day and temp
```{r}
df_ratio <- df_ratio[order(df_ratio$day,df_ratio$temp),]
```


Calculate log2 fold change
https://www.biostars.org/p/101727/

First calculate log2 fold change for 2 vs 3
```{r}

logFC_23 <- data.frame()
for (i in seq(2,nrow(df_ratio),3)){
  print(df_ratio[i,c(1:4)])
  print(df_ratio[i+ 1,c(1:4)])

  for(j in 1:ncol(df_ratio[,-c(1:4)])){
    logFC_23[i,j] <- log(df_ratio[i,j+4],2) - log(df_ratio[i+1,j+4],2)
    }
}
logFC_23 <- logFC_23[which(!is.na(logFC_23$V1)),]

logFC_23 <- cbind(unique(df_ratio[-1,"day"]), logFC_23)
colnames(logFC_23) <- c("day",colnames(df_ratio[,-c(1:4)]))


```

Next calculate log2 fold change for 9 vs 3
```{r}
logFC_93 <- data.frame()
for (i in seq(3,nrow(df_ratio),3)){
  print(df_ratio[i+ 1,c(1:4)])
  print(df_ratio[i,c(1:4)])

  for(j in 1:ncol(df_ratio[,-c(1:4)])){
    logFC_93[i,j] <- log(df_ratio[i+1,j+4],2) - log(df_ratio[i,j+4],2)
    }
}
logFC_93 <- logFC_93[which(!is.na(logFC_93$V1)),]

logFC_93 <- cbind(unique(df_ratio[-1,"day"]), logFC_93)
colnames(logFC_93) <- c("day",colnames(df_ratio[,-c(1:4)]))

```

First calculate log2 fold change for 9 vs 2
```{r}
logFC_92 <- data.frame()
for (i in seq(2,nrow(df_ratio),3)){
  print(df_ratio[i + 2,c(1:4)])
  print(df_ratio[i,c(1:4)])

  for(j in 1:ncol(df_ratio[,-c(1:4)])){
    logFC_92[i,j] <- log(df_ratio[i+2,j+4],2) - log(df_ratio[i,j+4],2)
    }
}
logFC_92 <- logFC_92[which(!is.na(logFC_92$V1)),]

logFC_92 <- cbind(unique(df_ratio[-1,"day"]), logFC_92)
colnames(logFC_92) <- c("day",colnames(df_ratio[,-c(1:4)]))

```


Calculate p value
https://www.r-bloggers.com/comparison-of-two-proportions-parametric-z-test-and-non-parametric-chi-squared-methods/
https://www.khanacademy.org/math/ap-statistics/two-sample-inference/two-sample-z-test-proportions/e/test-statistic-p-value-two-sample-z-test-proportions

*** I think x needs to be the numspectot for each protein, not the ratio. The ratio is only used to calculate the fold change


Calculate proportion test p value for 2 vs 3
```{r}
pval_23 <- data.frame()
for (i in seq(2,nrow(new_data_sum),3)){
  print(new_data_sum[i,c(1:4)])
  print(new_data_sum[i + 1,c(1:4)])
  print(spec_sum[i,])
  print(spec_sum[i+1,])
  for(j in 1:ncol(new_data_sum[,-c(1:4)])){
    obj <- prop.test(x = c(new_data_sum[i,j+4], new_data_sum[i+1,j+4]), n = c(spec_sum[i,"sum"],spec_sum[i+1,"sum"]), correct = FALSE) 
    pval_23[i,j] <- obj$p.value
    }
  }

###There were 50 or more warnings (use warnings() to see the first 50)
###Warnings()
###In prop.test(x = c(new_data_sum[1, j + 4], new_data_sum[i +  ... :
  ###Chi-squared approximation may be incorrect
pval_23 <- pval_23[which(!is.na(pval_23$V1)),]
pval_23 <- cbind(unique(new_data_sum[-1, "day"]), pval_23)
colnames(pval_23) <- c("day",colnames(new_data_sum[,-c(1:4)]))


#make a DF of FDR-adjusted pvals

pval_23_adj <- data.frame()
for (i in 1:nrow(pval_23)){
  padj <- p.adjust(pval_23[i,-1], method = "fdr")
  pval_23_adj <- rbind(pval_23_adj, padj)
}

pval_23_adj <- cbind(pval_23[,"day"], pval_23_adj)
colnames(pval_23_adj) <- colnames(pval_23)

```


Calculate proportion test p value for 9 vs 3
```{r}
pval_93 <- data.frame()
for (i in seq(3,nrow(new_data_sum),3)){
  print(new_data_sum[i+1,c(1:4)])
  print(new_data_sum[i,c(1:4)])
  print(spec_sum[i+1,])
  print(spec_sum[i,])
  for(j in 1:ncol(new_data_sum[,-c(1:4)])){
    obj <- prop.test(x = c(new_data_sum[i+1,j+4], new_data_sum[i,j+4]), n = c(spec_sum[i+1,"sum"],spec_sum[i,"sum"]), correct = FALSE) 
    pval_93[i,j] <- obj$p.value
    }
  }

###There were 50 or more warnings (use warnings() to see the first 50)
###Warnings()
###In prop.test(x = c(new_data_sum[1, j + 4], new_data_sum[i +  ... :
  ###Chi-squared approximation may be incorrect
pval_93 <- pval_93[which(!is.na(pval_93$V1)),]
pval_93 <- cbind(unique(new_data_sum[-1, "day"]), pval_93)
colnames(pval_93) <- c("day",colnames(new_data_sum[,-c(1:4)]))


#make a DF of FDR-adjusted pvals

pval_93_adj <- data.frame()
for (i in 1:nrow(pval_93)){
  padj <- p.adjust(pval_93[i,-1], method = "fdr")
  pval_93_adj <- rbind(pval_93_adj, padj)
}

pval_93_adj <- cbind(pval_93[,"day"], pval_93_adj)
colnames(pval_93_adj) <- colnames(pval_93)

```

Calculate proportion test p value for 9 vs 2
```{r}
pval_92 <- data.frame()
for (i in seq(2,nrow(new_data_sum),3)){
  print(new_data_sum[i + 2,c(1:4)])
  print(new_data_sum[i,c(1:4)])
  print(spec_sum[i+2,])
  print(spec_sum[i,])
  for(j in 1:ncol(new_data_sum[,-c(1:4)])){
    obj <- prop.test(x = c(new_data_sum[i+2,j+4], new_data_sum[i,j+4]), n = c(spec_sum[i+2,"sum"],spec_sum[i,"sum"]), correct = FALSE) 
    pval_92[i,j] <- obj$p.value
    }
  }

###There were 50 or more warnings (use warnings() to see the first 50)
###Warnings()
###In prop.test(x = c(new_data_sum[1, j + 4], new_data_sum[i +  ... :
  ###Chi-squared approximation may be incorrect
pval_92 <- pval_92[which(!is.na(pval_92$V1)),]
pval_92 <- cbind(unique(new_data_sum[-1, "day"]), pval_92)
colnames(pval_92) <- c("day",colnames(new_data_sum[,-c(1:4)]))


#make a DF of FDR-adjusted pvals

pval_92_adj <- data.frame()
for (i in 1:nrow(pval_92)){
  padj <- p.adjust(pval_92[i,-1], method = "fdr")
  pval_92_adj <- rbind(pval_92_adj, padj)
}

pval_92_adj <- cbind(pval_92[,"day"], pval_92_adj)
colnames(pval_92_adj) <- colnames(pval_92)

```

Join FC, logFC, pval, and pval_adj tables together

```{r}
#make a new column with sample calculation identifiers
logFC_23_SampleCalc <- paste("D",logFC_23$day, "logFC_23", sep = "_")
logFC_93_SampleCalc <- paste("D",logFC_93$day, "logFC_93", sep = "_")
logFC_92_SampleCalc <- paste("D",logFC_92$day, "logFC_92", sep = "_")

pval_23_SampleCalc <- paste("D",pval_23$day, "ChiSq.pval_23", sep = "_")
pval_93_SampleCalc <- paste("D",pval_93$day, "ChiSq.pval_93", sep = "_")
pval_92_SampleCalc <- paste("D",pval_92$day, "ChiSq.pval_92", sep = "_")

pval_23_adj_SampleCalc <- paste("D", pval_23_adj$day, "adj.ChiSq.pval_23", sep = "_")
pval_93_adj_SampleCalc <- paste("D", pval_93_adj$day, "adj.ChiSq.pval_93", sep = "_")
pval_92_adj_SampleCalc <- paste("D", pval_92_adj$day, "adj.ChiSq.pval_92", sep = "_")


tlogFC_23 <- t.data.frame(logFC_23[,-1])
colnames(tlogFC_23)<- logFC_23_SampleCalc

tlogFC_93 <- t.data.frame(logFC_93[,-1])
colnames(tlogFC_93)<- logFC_93_SampleCalc

tlogFC_92 <- t.data.frame(logFC_92[,-1])
colnames(tlogFC_92)<- logFC_92_SampleCalc

tpval_23 <- t.data.frame(pval_23[,-1])
colnames(tpval_23)<- pval_23_SampleCalc

tpval_93 <- t.data.frame(pval_93[,-1])
colnames(tpval_93)<- pval_93_SampleCalc

tpval_92 <- t.data.frame(pval_92[,-1])
colnames(tpval_92)<- pval_92_SampleCalc

tpval_23_adj <- t.data.frame(pval_23_adj[,-1])
colnames(tpval_23_adj) <- pval_23_adj_SampleCalc

tpval_93_adj <- t.data.frame(pval_93_adj[,-1])
colnames(tpval_93_adj) <- pval_93_adj_SampleCalc

tpval_92_adj <- t.data.frame(pval_92_adj[,-1])
colnames(tpval_92_adj) <- pval_92_adj_SampleCalc

logFC_pval_adjpval <- cbind(tlogFC_23,tpval_23,tpval_23_adj,tlogFC_93,tpval_93,tpval_93_adj,tlogFC_92,tpval_92,tpval_92_adj)

logFC_pval_adjpval <- logFC_pval_adjpval[,mixedsort(colnames(logFC_pval_adjpval))]

```




```{r, echo = FALSE, eval = FALSE}
write.csv(logFC_pval_adjpval, "~/Documents/GitHub/OysterSeedProject/analysis/TotNumSpecRatio_FC_Pval/sumNUMSPECSTOT_plus1_AllSilos_ratioFC_logFC_pval_DAYSCOMPARED.csv", quote = FALSE)

```


```{r}

What is the distribution of TOTNUMSPEC sums which aren't equal to 0?
```{r}
plot(hist(which(unlist(new_data_sum[,-c(1:4)]) !=0)))

length(which(unlist(new_data_sum[,-c(1:4)]) !=0 & unlist(new_data_sum[,-c(1:4)]) < 100))

plot(hist(which(unlist(new_data_sum[,-c(1:4)]) !=0 & < 100)))

```

What is the fraction of TOTNUMSPEC sums equal to 0
```{r}
length(which(unlist(new_data_sum[,-c(1:4)])==0))/length(unlist(new_data_sum[,-c(1:4)]))

```




```