---
title: "Untitled"
author: "Shelly Trigg"
date: "1/29/2019"
output: html_document
---

Load packages
```{r}
library(vegan)
library(ggplot2)
library(dplyr)
library(gtools)
```

Load Abacus data, parse out ADJNSAF values, and simplify column names to just sample number
```{r}
#upload data file
ABACUSdata <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/ABACUS_output021417.tsv", sep = "\t", header=TRUE, stringsAsFactors = FALSE)
#select only columns containing ADJNSAF and Protein ID
ABACUSdata <- ABACUSdata[,c(1,grep("NUMSPECSTOT", colnames(ABACUSdata)))]
ABACUSdata <- ABACUSdata[,-grep("ALL_", colnames(ABACUSdata))]

## change column names in ABACUSdata to just sampleID
colnames(ABACUSdata) <- gsub(pattern = "X20161205_SAMPLE_", "", colnames(ABACUSdata))
colnames(ABACUSdata) <- gsub(pattern = "_NUMSPECSTOT", "", colnames(ABACUSdata))
```

Load meta data file with temperature and day information
```{r}
#upload meta data; this was a csv file I create from Rhonda's notebook entry: https://github.com/Ellior2/Ellior2.github.io/blob/master/_posts/2017-3-11-NMDS-analysis.md
meta_data <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/Rhonda_new_sample_names.csv", header = TRUE, stringsAsFactors = FALSE)
meta_data$silo <- substr(meta_data$Contents,5,5)
meta_data$day <- substr(meta_data$SampleName,5,6)
meta_data$SampleName <- gsub(pattern = "H","",meta_data$SampleName)
meta_data$SampleName <- gsub(pattern = "C","",meta_data$SampleName)
#create a temperature column
meta_data$temp <- "temp"
for(i in 1:nrow(meta_data)){
  if(meta_data$silo[i] == "2"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "3"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "9"){
    meta_data$temp[i] <- "29"
  }
  if(meta_data$silo[i] == "e"){
    meta_data$temp[i] <- "16"
  }
}
```

Reformat Abacus data for NMDS
```{r}
#Transpose- switch rows and columns
tABACUSdata <- t.data.frame(ABACUSdata[,-1])
colnames(tABACUSdata) <- ABACUSdata[,1]
tABACUSdata <- cbind(data.frame(rownames(tABACUSdata)),tABACUSdata)
colnames(tABACUSdata)[1] <- "SampleID"

#add meta data to abacus data
tABACUSdata <- merge(meta_data[,c(1,2,7,8)],tABACUSdata, by = "SampleID")

#add meta data to abacus data
#Remove Silo 2 and day 15
silo3and9 <- tABACUSdata[which(substr(tABACUSdata$SampleName,1,2) != "S2" & tABACUSdata$day != "15"),]
#make rownames from Sample ID column so that the NMDS knows what's what
rownames(silo3and9) <- silo3and9$SampleID
#order the data frame by day and temperature so coloring the points on the plot is easier
silo3and9 <- silo3and9[order(as.numeric(silo3and9$day),silo3and9$temp),]

```

Determine if any proteins have zero NUMSPECSTOT vals for all samples; this would be because they were in Silo 2, but not in Silo 3 or 9
```{r}
no_val_proteins <- silo3and9[,which(apply(silo3and9, 2, var) == 0)]
ncol(no_val_proteins)
```

Remove proteins if they have a zero value in all samples
```{r}
silo3and9_nozerovar <- silo3and9[,-c(1:4,which(colnames(silo3and9) %in% colnames(no_val_proteins)))]
#check to make sure it worked
ncol(silo3and9)-ncol(silo3and9_nozerovar)
```

Sum NUMSPECSTOT for replicates
```{r}
df_sum <- data.frame()
#loop through the data and calculate the sum between replicates for each protein
for (i in seq(1,nrow(silo3and9_nozerovar),2)){
  #this calculates the sum for each odd number row and the row following it
  df_sum_row <- apply(silo3and9_nozerovar[c(i,i+1),],2,sum +1)
  #this sequencially combines rows of data together after the sum is generated
  df_sum <- rbind(df_sum, df_sum_row)
}
#add column names to sum'd data
colnames(df_sum) <- colnames(silo3and9_nozerovar)
df_sum$SampleID <- rownames(silo3and9_nozerovar[-grep("A",rownames(silo3and9_nozerovar)),])
new_data_sum <- merge(meta_data[,c("SampleID", "silo", "day", "temp")],df_sum, by = "SampleID")

```

remove proteins that had ADJNSAF value of 0 in all samples
```{r}
silo3and9_ADJNSAFnotzero <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/silo3and9_nozerovals.csv", stringsAsFactors = FALSE)

colnames(new_data_sum)<- gsub("\\|", ".", colnames(new_data_sum))

new_data_sum_ADJNSAFnotzero <- new_data_sum[,c(1:4,which(colnames(new_data_sum) %in% colnames(silo3and9_ADJNSAFnotzero[,-c(1:3)])))]
```

Add 1 to sum'd NUMSPECSTOT to eliminate zero values
```{r}
new_data_sum_ADJNSAFnotzero[,-c(1:4)] <- new_data_sum_ADJNSAFnotzero[,-c(1:4)] +1
```



Calculate sum of NUMSPECTOT for each sample
```{r}
spec_sum <- data.frame(apply(new_data_sum_ADJNSAFnotzero[,-c(1:4)], 1, sum))
spec_sum <- cbind(new_data_sum_ADJNSAFnotzero$SampleID,spec_sum)
colnames(spec_sum) <- c("SampleID", "sum")

```


Calculate ratio of Protein NUMSPECSTOT / Sample NUMSPECSTOT
```{r}
df_ratio <- data.frame()
#loop through the data and calculate the sum between replicates for each protein
for (i in 1:nrow(spec_sum)){
  #this calculates the sum for each odd number row and the row following it
  for(j in 1:ncol(new_data_sum_ADJNSAFnotzero[,-c(1:4)])){
    df_ratio[i,j] <- new_data_sum_ADJNSAFnotzero[i,j+4]/spec_sum[i,"sum"]*100000
  }  
}
#add column names to ratio data
colnames(df_ratio) <- colnames(new_data_sum_ADJNSAFnotzero[,-c(1:4)])
df_ratio <- cbind(new_data_sum_ADJNSAFnotzero[,1:4], df_ratio)
```


See what the lowest values are
```{r}
head(unique(sort(unlist(df_ratio[,-c(1:4)]))))
```

Calculate log2 fold change
https://www.biostars.org/p/101727/

```{r}
FC <- data.frame()
logFC <- data.frame()
for (i in 1:nrow(df_ratio[-1,])){
  for(j in 1:ncol(df_ratio[,-c(1:4)])){
    
    FC[i,j] <- (df_ratio[i+1,j+4] - df_ratio[1,j+4])/df_ratio[1,j+4]
    logFC[i,j] <- log(df_ratio[i+1,j+4],2) - log(df_ratio[1,j+4],2)
    }
}

FC <- cbind(df_ratio[-1, 1:4], FC)
colnames(FC) <- colnames(df_ratio)

logFC <- cbind(df_ratio[-1,1:4], logFC)
colnames(logFC) <- colnames(df_ratio)


```

Calculate p value
https://www.r-bloggers.com/comparison-of-two-proportions-parametric-z-test-and-non-parametric-chi-squared-methods/
https://www.khanacademy.org/math/ap-statistics/two-sample-inference/two-sample-z-test-proportions/e/test-statistic-p-value-two-sample-z-test-proportions

*** I think x needs to be the numspectot for each protein, not the ratio. The ratio is only used to calculate the fold change

```{r}
pval <- data.frame()
for (i in 1:nrow(new_data_sum_ADJNSAFnotzero[-1,])){
  for(j in 1:ncol(new_data_sum_ADJNSAFnotzero[,-c(1:4)])){
    obj <- prop.test(x = c(new_data_sum_ADJNSAFnotzero[1,j+4], new_data_sum_ADJNSAFnotzero[i+1,j+4]), n = c(spec_sum[1,"sum"],spec_sum[i+1,"sum"]), correct = FALSE) 
    pval[i,j] <- obj$p.value
    }
  }

###There were 50 or more warnings (use warnings() to see the first 50)
###Warnings()
###In prop.test(x = c(new_data_sum[1, j + 4], new_data_sum[i +  ... :
  ###Chi-squared approximation may be incorrect

pval <- cbind(new_data_sum_ADJNSAFnotzero[-1, 1:4], pval)
colnames(pval) <- colnames(new_data_sum_ADJNSAFnotzero)


#make a DF of FDR-adjusted pvals

pval_adj <- data.frame()
for (i in 1:nrow(pval)){
  padj <- p.adjust(pval[i,-c(1:4)], method = "fdr")
  pval_adj <- rbind(pval_adj, padj)
}

pval_adj <- cbind(new_data_sum_ADJNSAFnotzero[-1,1:4], pval_adj)
colnames(pval_adj) <- colnames(new_data_sum_ADJNSAFnotzero)

```

Join FC, logFC, pval, and pval_adj tables together

```{r}
#make a new column with sample calculation identifiers
FC_SampleCalc <- paste("D",FC$day, "T", FC$temp, "FC", sep = "_")
logFC_SampleCalc <- paste("D",logFC$day, "T", logFC$temp, "logFC", sep = "_")
pval_SampleCalc <- paste("D",pval$day, "T", pval$temp, "ChiSq.pval", sep = "_")
pval_adj_SampleCalc <- paste("D", pval_adj$day, "T", pval_adj$temp, "adj.ChiSq.pval", sep = "_")

tFC <- t.data.frame(FC[,-c(1:4)])
colnames(tFC)<- FC_SampleCalc

tlogFC <- t.data.frame(logFC[,-c(1:4)])
colnames(tlogFC)<- logFC_SampleCalc

tpval <- t.data.frame(pval[,-c(1:4)])
colnames(tpval)<- pval_SampleCalc

tpval_adj <- t.data.frame(pval_adj[,-c(1:4)])
colnames(tpval_adj) <- pval_adj_SampleCalc

FC_logFC_pval <- cbind(tFC,tlogFC,tpval, tpval_adj)

FC_logFC_pval <- FC_logFC_pval[,mixedsort(colnames(FC_logFC_pval))]

```

```{r, echo = FALSE, eval = FALSE}
write.csv(FC_logFC_pval, "~/Documents/GitHub/OysterSeedProject/analysis/TotNumSpecRatio_FC_Pval/sumNUMSPECSTOT_plus1_ratioFC_logFC_pval.csv", quote = FALSE)

```


```{r}
data <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/TotNumSpecRatio_FC_Pval/sumNUMSPECSTOT_plus1_ratioFC_logFC_pval.csv", stringsAsFactors = FALSE)

```