---
title: "Untitled"
author: "Shelly Trigg"
date: "1/29/2019"
output: html_document
---

Load packages
```{r}
library(vegan)
library(ggplot2)
library(dplyr)
library(gtools)
```

Load Abacus data, parse out ADJNSAF values, and simplify column names to just sample number
```{r}
#upload data file
ABACUSdata <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/ABACUS_output021417.tsv", sep = "\t", header=TRUE, stringsAsFactors = FALSE)
#select only columns containing ADJNSAF and Protein ID
ABACUSdata <- ABACUSdata[,c(1,grep("NUMSPECSTOT", colnames(ABACUSdata)))]
ABACUSdata <- ABACUSdata[,-grep("ALL_", colnames(ABACUSdata))]

## change column names in ABACUSdata to just sampleID
colnames(ABACUSdata) <- gsub(pattern = "X20161205_SAMPLE_", "", colnames(ABACUSdata))
colnames(ABACUSdata) <- gsub(pattern = "_NUMSPECSTOT", "", colnames(ABACUSdata))
```

Load meta data file with temperature and day information
```{r}
#upload meta data; this was a csv file I create from Rhonda's notebook entry: https://github.com/Ellior2/Ellior2.github.io/blob/master/_posts/2017-3-11-NMDS-analysis.md
meta_data <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/Rhonda_new_sample_names.csv", header = TRUE, stringsAsFactors = FALSE)
meta_data$silo <- substr(meta_data$Contents,5,5)
meta_data$day <- substr(meta_data$SampleName,5,6)
meta_data$SampleName <- gsub(pattern = "H","",meta_data$SampleName)
meta_data$SampleName <- gsub(pattern = "C","",meta_data$SampleName)
#create a temperature column
meta_data$temp <- "temp"
for(i in 1:nrow(meta_data)){
  if(meta_data$silo[i] == "2"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "3"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "9"){
    meta_data$temp[i] <- "29"
  }
  if(meta_data$silo[i] == "e"){
    meta_data$temp[i] <- "16"
  }
}
```

Reformat Abacus data for NMDS
```{r}
#Transpose- switch rows and columns
tABACUSdata <- t.data.frame(ABACUSdata[,-1])
colnames(tABACUSdata) <- ABACUSdata[,1]
tABACUSdata <- cbind(data.frame(rownames(tABACUSdata)),tABACUSdata)
colnames(tABACUSdata)[1] <- "SampleID"

#add meta data to abacus data
tABACUSdata <- merge(meta_data[,c(1,2,7,8)],tABACUSdata, by = "SampleID")

#add meta data to abacus data
#Remove Silo 2 and day 15
silo3and9 <- tABACUSdata[which(substr(tABACUSdata$SampleName,1,2) != "S2" & tABACUSdata$day != "15"),]
#make rownames from Sample ID column so that the NMDS knows what's what
rownames(silo3and9) <- silo3and9$SampleID
#order the data frame by day and temperature so coloring the points on the plot is easier
silo3and9 <- silo3and9[order(as.numeric(silo3and9$day),silo3and9$temp),]

```

Determine if any proteins have zero ADJNSAF vals for all samples; this would be because they were in Silo 2, but not in Silo 3 or 9
```{r}
no_val_proteins <- silo3and9[,which(apply(silo3and9, 2, var) == 0)]
ncol(no_val_proteins)
```

Remove proteins if they have a zero value in all samples
```{r}
silo3and9_nozerovar <- silo3and9[,-c(1:4,which(colnames(silo3and9) %in% colnames(no_val_proteins)))]
#check to make sure it worked
ncol(silo3and9)-ncol(silo3and9_nozerovar)
```

Sum NUMSPECSTOT for replicates
```{r}
df_sum <- data.frame()
#loop through the data and calculate the sum between replicates for each protein
for (i in seq(1,nrow(silo3and9_nozerovar),2)){
  #this calculates the sum for each odd number row and the row following it
  df_sum_row <- apply(silo3and9_nozerovar[c(i,i+1),],2,sum)
  #this sequencially combines rows of data together after the sum is generated
  df_sum <- rbind(df_sum, df_sum_row)
}
#add column names to SD data
colnames(df_sum) <- colnames(silo3and9_nozerovar)
df_sum$SampleID <- rownames(silo3and9_nozerovar[-grep("A",rownames(silo3and9_nozerovar)),])
new_data_sum <- merge(meta_data[,c("SampleID", "silo", "day", "temp")],df_sum, by = "SampleID")

```

Calculate sum of NUMSPECTOT for each sample
```{r}
spec_sum <- data.frame(apply(new_data_sum[,-c(1:4)], 1, sum))
colnames(spec_sum)[1] <- "sum"
```


Calculate ratio of Protein NUMSPECSTOT / Sample NUMSPECSTOT
```{r}
df_ratio <- data.frame()
#loop through the data and calculate the sum between replicates for each protein
for (i in 1:nrow(spec_sum)){
  #this calculates the sum for each odd number row and the row following it
  for(j in 1:ncol(new_data_sum[,-c(1:4)])){
    df_ratio[i,j] <- new_data_sum[i,j+4]/spec_sum[i,"sum"]*100000
  }  
}
#add column names to SD data
colnames(df_ratio) <- colnames(new_data_sum[,-c(1:4)])
df_ratio <- cbind(new_data_sum[,1:4], df_ratio)
```

remove proteins that had ADJNSAF value of 0 in all samples
```{r}
silo3and9_ADJNSAFnotzero <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/silo3and9_nozerovals.csv", stringsAsFactors = FALSE)

colnames(df_ratio)<- gsub("\\|", ".", colnames(df_ratio))

df_ratio_ADJNSAFnotzero <- df_ratio[,c(1:4,which(colnames(df_ratio) %in% colnames(silo3and9_ADJNSAFnotzero[,-c(1:3)])))]
```


For proteins with a zero value in any sample, replace with very small value
```{r}
df_ratio_ADJNSAFnotzero[df_ratio_ADJNSAFnotzero == 0] <- 0.01000
```

Calculate log2 fold change
https://www.biostars.org/p/101727/
```{r}
S3D3 <- data.frame(t.data.frame(df_ratio_ADJNSAFnotzero[which(df_ratio_ADJNSAFnotzero$SampleID == "1" | df_ratio_ADJNSAFnotzero$SampleID == "4"),-c(1:4)]))
colnames(S3D3) <- c(paste(df_ratio_ADJNSAFnotzero$silo[1],df_ratio_ADJNSAFnotzero$day[1], sep = "_"),paste(df_ratio_ADJNSAFnotzero$silo[9],df_ratio_ADJNSAFnotzero$day[9], sep = "_"))
  
S3D3$FC <- (S3D3$`3_3`- S3D3$e_0.01)/S3D3$e_0.01
S3D3$logFC <- log(S3D3$`3_3`,2) - log(S3D3$e_0.01,2)
```

Calculate p value
https://www.r-bloggers.com/comparison-of-two-proportions-parametric-z-test-and-non-parametric-chi-squared-methods/
https://www.khanacademy.org/math/ap-statistics/two-sample-inference/two-sample-z-test-proportions/e/test-statistic-p-value-two-sample-z-test-proportions

```{r}
pval <- data.frame()
for (i in 1:nrow(S3D3)){
obj <- prop.test(x = c(S3D3$e_0.01[i], S3D3$`3_3`[i]), n = c(108123,112194), correct = FALSE) 
pval[i,1] <- obj$p.value
}

S3D3 <- cbind(S3D3, pval)
colnames(S3D3)[5] <- "ChiSq_p.val"
```

