---
title: "Untitled"
author: "Shelly Trigg"
date: "9/16/2019"
output: html_document
---

###############################################################################
###############################################################################
############Many failed analysis attempts#################################
###############################################################################
###############################################################################

###run in ASCA_logNSAFs_FunctionalAnalysis.Rproj environment (many objects created in other scripts in the environment)


**Run a proportions test to see if proportions of terms in sig protein list are different than their proportions in the background list**
  
  
  ```{r}
#first make background with same terms that are in each clade list


clade1_bkgd <- STACKED_silo23_ALL_pro_GOid_term[which(STACKED_silo23_ALL_pro_GOid_term$GO %in% silo23_clade1_GO$GO),]

clade1_bkgd_freqs <- data.frame(table(clade1_bkgd$GO))
clade1_bkgd_total_terms <- nrow(clade1_bkgd)

clade1_sig_freqs <- data.frame(table(silo23_clade1_GO$GO))
clade1_total_terms <- nrow(silo23_clade1_GO)

clade1_term_obsv <- merge(clade1_sig_freqs,clade1_bkgd_freqs, by = "Var1")
colnames(clade1_term_obsv) <- c("GO_ID","Sig.Freq", "Bkgd.Freq")

for (i in (1:nrow(clade1_term_obsv))){
  prop.results <- prop.test(x=c(clade1_term_obsv[i,"Sig.Freq"],clade1_term_obsv[i,"Bkgd.Freq"]), n=c(clade1_total_terms,clade1_bkgd_total_terms))
  clade1_term_obsv$p.value[i] <- prop.results$p.value
  clade1_term_obsv$p.adj[i] <- p.adjust(clade1_term_obsv$p.value[i], method = "fdr", n = nrow(clade1_term_obsv))
  
}


```


**GO MWU**
  ```{r}

#save proteins and GO_IDs for GO MWU input
silo23_ALL_pro_GO_mwu <- cbind(silo23_ALL_pro_GO[,1],data.frame(gsub(" ","",silo23_ALL_pro_GO$GO_IDs)))
write.table(silo23_ALL_pro_GO_mwu,"~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo23_all_pro_GO.tab",quote = FALSE, sep = "\t", col.names = FALSE,row.names = FALSE)


#edit measure of significance table for passing it to MWU
mwu_table <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo23_ASCA_PC1_siloeffect.csv", stringsAsFactors = FALSE)
#probably not necessary to match column names between input files
colnames(mwu_table)[1] <- "protein_ID"
#probably not necessary: remove the pipe in the protein names
mwu_table$protein_ID <- gsub("\\|","\\.",mwu_table$protein_ID)


#convert PC1loadings to absolute values to that ranking works
mwu_table$PC1loadings<- abs(mwu_table$PC1loadings)

write.csv(mwu_table,"~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo23_ASCA_Prot_PC1_siloeffect.csv", quote = FALSE, row.names = FALSE)

```



```{r}
# Edit these to match your data file names: 
input="silo23_ASCA_Prot_PC1_siloeffect.csv" # two columns of comma-separated values: gene id, continuous measure of significance. To perform standard GO enrichment analysis based on Fisher's exact test, use binary measure (0 or 1, i.e., either sgnificant or not).
goAnnotations="silo23_all_pro_GO.tab" # two-column, tab-delimited, one line per gene, multiple GO terms separated by semicolon. If you have multiple lines per gene, use nrify_GOtable.pl prior to running this script.
goDatabase="go.obo" # download from http://www.geneontology.org/GO.downloads.ontology.shtml
goDivision="BP" # either MF, or BP, or CC
source("gomwu.functions.R")

```


```{r}
# Calculating stats. It might take ~3 min for MF and BP. Do not rerun it if you just want to replot the data with different cutoffs, go straight to gomwuPlot. If you change any of the numeric values below, delete the files that were generated in previos runs first.
gomwuStats(input, goDatabase, goAnnotations, goDivision,
           perlPath="perl", # replace with full path to perl executable if it is not in your system's PATH already
           largest=0.1,  # a GO category will not be considered if it contains more than this fraction of the total number of genes
           smallest=5,   # a GO category should contain at least this many genes to be considered
           clusterCutHeight=0.25) # threshold for merging similar (gene-sharing) terms. See README for details.
# There are no GO terms pass 10% FDR.
```

```{r}
# Plotting results
quartz()
results=gomwuPlot(input,goAnnotations,goDivision,
                  #	absValue=-log(0.05,10),  # genes with the measure value exceeding this will be counted as "good genes". Specify absValue=0.001 if you are doing Fisher's exact test for standard GO enrichment or analyzing a WGCNA module (all non-zero genes = "good genes").
                  absValue=1,
                  level1=0.1, # FDR threshold for plotting. Specify level1=1 to plot all GO categories containing genes exceeding the absValue.
                  level2=0.05, # FDR cutoff to print in regular (not italic) font.
                  level3=0.01, # FDR cutoff to print in large bold font.
                  txtsize=1.2,    # decrease to fit more on one page, or increase (after rescaling the plot so the tree fits the text) for better "word cloud" effect
                  treeHeight=0.5, # height of the hierarchical clustering tree
                  #	colors=c("dodgerblue2","firebrick1","skyblue2","lightcoral") # these are default colors, un-remar and change if needed
)
# manually rescale the plot so the tree matches the text 
# if there are too many categories displayed, try make it more stringent with level1=0.05,level2=0.01,level3=0.001.  

# text representation of results, with actual adjusted p-values
results

```

#read in results
```{r}
silo23_GOmwu_results <- read.table("~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/MWU_BP_silo23_ASCA_Prot_PC1_siloeffect.csv", sep = " ", header = TRUE, stringsAsFactors = FALSE)

```











**Map all GO IDs to GO slim terms; not sure how helpful this is **
  
  ```{r}
#make list of unique GO terms without a protein ID column
silo23_sig_GOids <- unique(STACKED_silo23_sig_pro_GOid_term$GO)
#Use GSEA to generate list of all GO Slim BP, MP, and CC
#DF will have "term", "GOid", "GOcategory"

#BP first
#goslims with GSEA
myCollection <- GOCollection(silo23_sig_GOids)
#I downloaded goslim_generic.obo from http://geneontology.org/docs/go-subset-guide/
#then i moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims <- data.frame(goSlim(myCollection, slim, "BP"))
slims$GOid <- rownames(slims)
slims$Term <- as.character(slims$Term)
rownames(slims) <- NULL

GSEA_bp <- slims[,c("Term", "GOid")]
GSEA_bp$GOcategory <- "BP"

#MP next
slims <- data.frame(goSlim(myCollection, slim, "MF"))
slims$GOid <- rownames(slims)
slims$Term <- as.character(slims$Term)
rownames(slims) <- NULL

GSEA_MF <- slims[,c("Term", "GOid")]
GSEA_MF$GOcategory <- "MF"

GSEA_BP_MF <- rbind(GSEA_bp, GSEA_MF)
colnames(GSEA_BP_MF) <- c("Term", "GOslim", "GOcategory")
```


**make a list of GO slims for background proteins**
  ```{r}
#make list of unique GO terms without a protein ID column
silo23_ALL_GOids <- unique(STACKED_silo23_ALL_pro_GOid_term$GO)
#Use GSEA to generate list of all GO Slim BP, MP, and CC
#DF will have "term", "GOid", "GOcategory"

#BP first
myCollection.All <- GOCollection(silo23_ALL_GOids)

slims_ALL <- data.frame(goSlim(myCollection.All, slim, "BP"))
slims_ALL$GOid <- rownames(slims_ALL)
slims$Term <- as.character(slims$Term)
rownames(slims) <- NULL

GSEA_bp <- slims[,c("Term", "GOid")]
GSEA_bp$GOcategory <- "BP"



#tried prop test but nothing significant
# > sum(slims$Count)
# [1] 1081
# > sum(slims_ALL$Count)
# [1] 19170
# > test <- prop.test(x=c(29,413), n = c(1081,19170))
# > test
# 
# 	2-sample test for equality of proportions with
# 	continuity correction
# 
# data:  c(29, 413) out of c(1081, 19170)
# X-squared = 1.1017, df = 1, p-value = 0.2939
# alternative hypothesis: two.sided
# 95 percent confidence interval:
#  -0.005054521  0.015620387
# sample estimates:
#     prop 1     prop 2 
# 0.02682701 0.02154408 
# > test <- prop.test(x=c(23,390), n = c(1081,19170))
# > test
# 
# 	2-sample test for equality of proportions with
# 	continuity correction
# 
# data:  c(23, 390) out of c(1081, 19170)
# X-squared = 0.010083, df = 1, p-value = 0.92
# alternative hypothesis: two.sided
# 95 percent confidence interval:
#  -0.00838773  0.01025235
# sample estimates:
#     prop 1     prop 2 
# 0.02127660 0.02034429 


```

**TRY FOLD CHANGE ANALYSIS**
  ```{r}
silo23_ASCA_FC <- silo23_ASCA
silo23_ASCA_FC$silo2_avg <- apply(silo23_ASCA_FC[,which(substr(colnames(silo23_ASCA_FC),2,2)==2)],1,mean)
silo23_ASCA_FC$silo3_avg <- apply(silo23_ASCA_FC[,which(substr(colnames(silo23_ASCA_FC),2,2)==3)],1,mean)


```



**TRY ONTOLOGY X packages**
  ```{r}

data(go)


###customizing gene list to use
agg_df <- aggregate(GO~protein_ID, silo23_clade1_GO, c)
agg_list <- as.list(agg_df$GO)
names(agg_list)<- agg_df$protein_ID

bp <- go$id[go$name == "biological_process"]
bp_anno <- lapply(agg_list,function(x) get_ancestors(go, intersection_with_descendants(go, roots=bp, x)))
all_terms <- unique(unlist(use.names=FALSE, bp_anno))
terms <- remove_uninformative_terms(go, agg_list)

rmlinks_terms <- remove_links(go,terms= terms)

length(all_terms)
length(terms)
length(rmlinks_terms)

###still too many terms (793). Can't plot this because it's too messy. It needs to be further reduced



###this didn't get anywhere
#extract unique GO terms for clade 1
silo23_clade1_GO_uniq <- unique(silo23_clade1_GO$GO)

#get parent terms of unique GO IDs 
silo23_clade1_GO_parents<- list()
for(i in 1:length(silo23_clade1_GO_uniq)){
  temp_silo23_clade1_GO_parents <- try(go$parents[[silo23_clade1_GO_uniq[i]]], TRUE)
  if(isTRUE(class(temp_silo23_clade1_GO_parents)=="try-error")) {next} else {silo23_clade1_GO_parents[[i]] = temp_silo23_clade1_GO_parents}
}

names(silo23_clade1_GO_parents) <- silo23_clade1_GO_uniq

#silo23_clade1_GO_parents <- silo23_clade1_GO_parents[!sapply(silo23_clade1_GO_parents, is.null)]

silo23_clade1_GO_index <- ontology_index(parents = silo23_clade1_GO_parents)

#remove redundant terms
silo23_clade1_GO_index_min <- minimal_set(silo23_clade1_GO_index,silo23_clade1_GO_uniq)

silo23_clade1_GO_index_min_inform <- remove_uninformative_terms(go,silo23_clade1_GO_index_min)

bp <- go$id[go$name == "biological_process"]
bp_anno <- lapply(silo23_clade1_GO_index_min,function(x) get_ancestors(go, intersection_with_descendants(go, roots=bp, x)))

onto_plot(go,terms = get_ancestors(go,silo23_clade1_GO_index_min))

terms <- remove_links(go, get_ancestors(go, silo23_clade1_GO_index_min))

onto_plot(go, terms=terms, fillcolor=rainbow(length(terms)))

```



**map to SR lab GO slims**
```{r}
srlabslim <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/background/GOSlim_terms.csv", stringsAsFactors = FALSE)
colnames(srlabslim)[1] <- "GO"
silo23_clade1_GO_srlabslim <- merge(silo23_clade1_GO, srlabslim, by = "GO")
View(table(silo23_clade1_GO_srlabslim$GOSlim_bin))
#results are too general and not informative, nothing to compare to, not screaming out that anything is happening 
```


**Create Ontology Plots**
```{r}
library(ontologyIndex)
library(ontologyPlot)
library(ontologySimilarity)
data(go)
data(gene_GO_terms)
data(GO_IC)
cc <- go$id[go$name == "biological_process"]
cc_anno <- lapply(gene_GO_terms[c("QPCTL","CRNN")],function(x) get_ancestors(go, intersection_with_descendants(go, roots=cc, x)))
all_terms <- unique(unlist(use.names=FALSE, cc_anno))
QPCTL <- cc_anno$QPCTL
CRNN <- cc_anno$CRNN
terms <- remove_uninformative_terms(go, list(QPCTL, CRNN))
pdf(file="GO-plot.pdf", width=5, height=4)
onto_plot(go, terms=terms,width=calibrate_sizes(GO_IC[terms], low=1, high=3),fillcolor=rgb(0, terms %in% QPCTL, terms %in% CRNN, 0.5))
dev.off()



###customizing gene list to use
agg_df <- aggregate(GO~protein_ID, silo23_clade1_GO, c)
agg_list <- as.list(agg_df$GO)
names(agg_list)<- agg_df$protein_ID

bp <- go$id[go$name == "biological_process"]
bp_anno <- lapply(agg_list,function(x) get_ancestors(go, intersection_with_descendants(go, roots=bp, x)))
all_terms <- unique(unlist(use.names=FALSE, bp_anno))
terms <- remove_uninformative_terms(go, agg_list)

rmlinks_terms <- remove_links(go,terms= terms)

length(all_terms)
length(terms)
length(rmlinks_terms)

pdf(file="GO-plot.pdf", width=5, height=4)
onto_plot(go, terms=terms,width=calibrate_sizes(GO_IC[terms], low=1, high=3),fillcolor=rgb(0, terms %in% QPCTL, terms %in% CRNN, 0.5))
dev.off()


**remove GO terms with sem sim values < 0.5 and !=1, then try heatmap**
```{r}
sim_matrix_m <- melt(sim_matrix)
sim_matrix_m <- sim_matrix_m[which(sim_matrix_m$value>0.5),]
sim_matrix_m <- sim_matrix_m[which(sim_matrix_m$value!=1),]
sim_matrix_m <- tidyr::spread(sim_matrix_m, "Var2","value")
rownames(sim_matrix_m) <- sim_matrix_m$Var1
sim_matrix_m <- sim_matrix_m[,-1]
sim_matrix_m[is.na(sim_matrix_m)] <- 0
sim_matrix_m_sd <- data.frame(apply(sim_matrix_m,1,sd))
colnames(sim_matrix_m_sd) <- "SD"

sim_matrix_m<- sim_matrix_m[which(!(rownames(sim_matrix_m)%in% rownames(sim_matrix_m_sd[which(sim_matrix_m_sd$SD > 0),]))),]


### Always had too many terms so never got a clear plot
```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)
library(topGO)


#background list silo39_ALL_pro_GO
#sig pro list silo39_sig_pro_GO

topgo_bkgd <- silo39_ALL_pro_GO
topgo_bkgd$GO_IDs <- gsub(";",",",topgo_bkgd$GO_IDs)


#to switch between time and temp run line 66 in time and temp functional analysis scripts
topgo_sig <- silo39_sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterID == 1),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterID == 2),c("protein_ID", "GO_IDs")]

topgo_sig_3 <- topgo_sig[which(topgo_sig$ClusterID == 3),c("protein_ID", "GO_IDs")]

topgo_sig_4 <- topgo_sig[which(topgo_sig$ClusterID == 4),c("protein_ID", "GO_IDs")]

topgo_sig_5 <- topgo_sig[which(topgo_sig$ClusterID == 5),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="time_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="time_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 3

genesOfInterest_3 <- topgo_sig_3
genesOfInterest_3 <- as.character(genesOfInterest_3$protein_ID)

geneList_3 <- factor(as.integer(geneUniverse %in% genesOfInterest_3))
names(geneList_3) <- geneUniverse

myGOdata_3 <- new("topGOdata", description="time_clade3", ontology="BP", allGenes=geneList_3,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_3 <- runTest(myGOdata_3, algorithm="weight01", statistic="fisher") 

allRes_3 <- GenTable(myGOdata_3, weight = resultFisher_3,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 4

genesOfInterest_4 <- topgo_sig_4
genesOfInterest_4 <- as.character(genesOfInterest_4$protein_ID)

geneList_4 <- factor(as.integer(geneUniverse %in% genesOfInterest_4))
names(geneList_4) <- geneUniverse

myGOdata_4 <- new("topGOdata", description="time_clade4", ontology="BP", allGenes=geneList_4,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_4 <- runTest(myGOdata_4, algorithm="weight01", statistic="fisher") 

allRes_4 <- GenTable(myGOdata_4, weight = resultFisher_4,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 5

genesOfInterest_5 <- topgo_sig_5
genesOfInterest_5 <- as.character(genesOfInterest_5$protein_ID)

geneList_5 <- factor(as.integer(geneUniverse %in% genesOfInterest_5))
names(geneList_5) <- geneUniverse

myGOdata_5 <- new("topGOdata", description="time_clade5", ontology="BP", allGenes=geneList_5,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_5 <- runTest(myGOdata_5, algorithm="weight01", statistic="fisher") 


allRes_5 <- GenTable(myGOdata_5, weight = resultFisher_5,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]
sub_allRes3 <- allRes_3[which(allRes_3$Annotated >=5 & allRes_3$weight <= 0.05),]
sub_allRes4 <- allRes_4[which(allRes_4$Annotated >=5 & allRes_4$weight <= 0.05),]
sub_allRes5 <- allRes_5[which(allRes_5$Annotated >=5 & allRes_5$weight <= 0.05),]


#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2,sub_allRes3,sub_allRes4,sub_allRes5)

sub_allRes <- rbind(sub_allRes1, sub_allRes2)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

sub_allRes <- merge(sub_allRes, sub_allRes3[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[5] <- "clade3"

sub_allRes <- merge(sub_allRes, sub_allRes4[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[6] <- "clade4"

sub_allRes <- merge(sub_allRes, sub_allRes5[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[7] <- "clade5"


```

#get GO slims for clade terms
```{r}
library(GSEABase)
#make list of unique GO terms without a protein ID column
GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]
GO_clade3 <- sub_allRes[which(!is.na(sub_allRes$clade3)),"GO.ID"]
GO_clade4 <- sub_allRes[which(!is.na(sub_allRes$clade4)),"GO.ID"]
GO_clade5 <- sub_allRes[which(!is.na(sub_allRes$clade5)),"GO.ID"]



#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)
collection_clade3 <- GOCollection(GO_clade3)
collection_clade4 <- GOCollection(GO_clade4)
collection_clade5 <- GOCollection(GO_clade5)

#I downloaded goslim_generic.obo from http://geneontology.org/docs/go-subset-guide/
#then i moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))
slims_clade3 <- data.frame(goSlim(collection_clade3, slim, "BP"))
slims_clade4 <- data.frame(goSlim(collection_clade4, slim, "BP"))
slims_clade5 <- data.frame(goSlim(collection_clade5, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2, slims_clade3, slims_clade4, slims_clade5)

slims <- rbind(slims_clade1, slims_clade2)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

slims <- merge(slims, slims_clade3[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[4] <- "clade3"

slims <- merge(slims, slims_clade4[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[5] <- "clade4"

slims <- merge(slims, slims_clade5[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[6] <- "clade5"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]

slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))




#library(viridis)
#col = rev(viridis(11))
col = c("#FFFFFFFF","#C2DF23FF","#85D54AFF","#51C56AFF","#2BB07FFF","#1E9B8AFF","#25858EFF","#2D708EFF","#38598CFF","#433E85FF","#482173FF","#440154FF")
breaks <- c(-1,0,1,2,3,4,5,6,7,8,9,10,11)
#heatmap.2(slims_m,margins = c(5,15),cexRow = 0.5, cexCol = 0.75, Colv=NA, hclustfun=function(x) rev(as.dendrogram(hclust(x, method="complete"))),breaks= breaks,col = col,density.info = "none", trace = "none")
#pdf(file="GOSlimheatmapTime.pdf", width=8, height=8)
#heatmap.2(slims_m,margins = c(5,20),cexRow = 0.8, Colv=NA,revC = TRUE,breaks= breaks,col = col,density.info = "none", trace = "none", scale = "none")
#dev.off()

ColSideColors<- c("#000000","#00FF00","#808080","#8000FF","#80FFFF")
pdf(file="GOSlimheatmapTime.pdf", width=8, height=8)
heatmap.2(slims_m,margins = c(5,20),cexRow = 0.8, Colv=NA,revC = TRUE,breaks= breaks,col = col,density.info = "none", trace = "none", scale = "none",ColSideColors = ColSideColors)
dev.off()

ColSideColors<- c("#000000","#80FFFF")
pdf(file="GOSlimheatmapTemp.pdf", width=8, height=8)
heatmap.2(slims_m,margins = c(5,35),cexCol = 1, Colv=NA,revC = TRUE,breaks= breaks,col = col,density.info = "none", trace = "none", scale = "none",ColSideColors = ColSideColors)
dev.off()

```