---
title: "Silo3_9_ASCA_logNSAFs_FunctionalAnalysis"
author: "Shelly Trigg"
date: "9/14/2019"
output: html_document
---

**load libraries**
```{r}
#load libraries
library(plyr)
library(tidyr)
library(GSEABase)
library(reshape2)
library(gplots)
library(viridis)
library(topGO)
```

**read in and format uniprot data**
```{r}
#read in uniprot mapping
uniprot <- read.csv("/Volumes/web/metacarcinus/Cgigas/all_giga-uniprot-blastP-out.nopipe.annotations.tab", sep ="\t", header = FALSE, stringsAsFactors = FALSE)
#rename uniprot columns
colnames(uniprot) <- c("protein_ID","Entry", "Entry_name", "perc_ident_match", "align_len", "num_mismatch", "num_gaps","querStart", "querEnd", "subjStart", "subjEnd", "evalue", "bitscore","Entry.1","Entry_name.1", "Protein_names", "Gene_names", "Organism", "Protein_length","Pathway", "GO_bp", "GO","GO_IDs", "Protein_fams")
#convert columns that should be numeric to numeric
str(uniprot)
uniprot[,4:13] <- lapply(uniprot[,4:13], as.numeric)
#NAs introduced by coercion
#this is because the entries that are unmappaed have NA values
str(uniprot)

#filter for choyp proteins and for evalue <= 10x10^-10 (unmapped proteins get removed here)
uniprot_filtered <- uniprot[grep("CHOYP",uniprot$protein_ID),]
uniprot_filtered <- uniprot[which(uniprot$evalue <= 1*10^-10),]


```

**read in and format ASCA logNSAF time and temp sig. data**
```{r}
silo39_time_ASCA <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/ASCA/ASCA_all_proteins_avgADJNSAF/Silo3_9_log/silo3_9_timeSig_prots.csv", stringsAsFactors = FALSE)

```

**read in all proteins detected specific for each comparison to generate GO background 
```{r}
silo39_all_prots <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/ASCA/ASCA_all_proteins_avgADJNSAF/Silo3_9_log/all_silo39_prots.csv", stringsAsFactors = FALSE)

#rename column to protein_ID
colnames(silo39_all_prots)[1] <- "protein_ID"
#remove pipe in protein IDs
silo39_all_prots$protein_ID <- gsub("\\|","\\.", silo39_all_prots$protein_ID)


```

**Match GO terms to sig protein lists**

```{r}
#Get GO IDs for silo 3 and 9 sig. proteins
silo39_sig_pro_GO <- merge(silo39_time_ASCA, uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(silo39_sig_pro_GO)

#Since GO IDs are listed all in one column, spread them across multiple columns (1 column /term)
silo39_sig_pro_GOid_term <- data.frame()
for (i in 1:nrow(silo39_sig_pro_GO)){
  silo39_sig_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo39_sig_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))))
  silo39_sig_pro_GOid_term <- rbind.fill(silo39_sig_pro_GOid_term,silo39_sig_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo39_sig_pro_GOid_term <- cbind(silo39_sig_pro_GO[,"protein_ID"], silo39_sig_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
#convert all factors to  characters
silo39_sig_pro_GOid_term <- data.frame(lapply(silo39_sig_pro_GOid_term, as.character), stringsAsFactors = FALSE)
#proteins that don't have GO IDs?

#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo39_sig_pro_GOid_term <- tidyr::gather(silo39_sig_pro_GOid_term,"col","GO", 2:ncol(silo39_sig_pro_GOid_term))
#exlude middle column which just contains the string "gene" in each row
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[,-2]
#remove duplicate rows
STACKED_silo39_sig_pro_GOid_term <- unique(STACKED_silo39_sig_pro_GOid_term)
colnames(STACKED_silo39_sig_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[which(!is.na(STACKED_silo39_sig_pro_GOid_term$GO)),]
#this resulting data frame has two columns "gene" and "GO"
#write.csv(STACKED_sig0.1_pro_GOid_term, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/SameDayFCtoTemp/ChiSq_ASCA_clust/intermediate_files/STACKED_sig0.1_pro_GOid_term.csv", quote = FALSE, row.names = FALSE)
```

**MATCH GO TERMS TO BACKGROUND LIST**
This step takes ~10min
```{r}
#Get GO IDs for silo 3 and 9 ALL. proteins
silo39_ALL_pro_GO <- merge(silo39_all_prots, uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(silo39_ALL_pro_GO)

#Since GO IDs are listed all in one column, spread them across multiple columns (1 column /term)
silo39_ALL_pro_GOid_term <- data.frame()
for (i in 1:nrow(silo39_ALL_pro_GO)){
  silo39_ALL_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo39_ALL_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))))
  silo39_ALL_pro_GOid_term <- rbind.fill(silo39_ALL_pro_GOid_term,silo39_ALL_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo39_ALL_pro_GOid_term <- cbind(silo39_ALL_pro_GO[,"protein_ID"], silo39_ALL_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
#convert all factors to  characters
silo39_ALL_pro_GOid_term <- data.frame(lapply(silo39_ALL_pro_GOid_term, as.character), stringsAsFactors = FALSE)
#proteins that don't have GO IDs?

#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo39_ALL_pro_GOid_term <- tidyr::gather(silo39_ALL_pro_GOid_term,"col","GO", 2:ncol(silo39_ALL_pro_GOid_term))
#exlude middle column which just contains the string "gene" in each row
STACKED_silo39_ALL_pro_GOid_term <- STACKED_silo39_ALL_pro_GOid_term[,-2]
#remove duplicate rows
STACKED_silo39_ALL_pro_GOid_term <- unique(STACKED_silo39_ALL_pro_GOid_term)
colnames(STACKED_silo39_ALL_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo39_ALL_pro_GOid_term <- STACKED_silo39_ALL_pro_GOid_term[which(!is.na(STACKED_silo39_ALL_pro_GOid_term$GO)),]
#this resulting data frame has two columns "gene" and "GO"
#write.csv(STACKED_ALL0.1_pro_GOid_term, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/SameDayFCtoTemp/ChiSq_ASCA_clust/intermediate_files/STACKED_ALL0.1_pro_GOid_term.csv", quote = FALSE, row.names = FALSE)

```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)
library(topGO)


#background list silo39_ALL_pro_GO
#sig pro list silo39_sig_pro_GO

topgo_bkgd <- silo39_ALL_pro_GO
topgo_bkgd$GO_IDs <- gsub(";",",",topgo_bkgd$GO_IDs)


#to switch between time and temp run line 66 in time and temp functional analysis scripts
topgo_sig <- silo39_sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterID == 1),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterID == 2),c("protein_ID", "GO_IDs")]

topgo_sig_3 <- topgo_sig[which(topgo_sig$ClusterID == 3),c("protein_ID", "GO_IDs")]

topgo_sig_4 <- topgo_sig[which(topgo_sig$ClusterID == 4),c("protein_ID", "GO_IDs")]

topgo_sig_5 <- topgo_sig[which(topgo_sig$ClusterID == 5),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="time_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="time_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 3

genesOfInterest_3 <- topgo_sig_3
genesOfInterest_3 <- as.character(genesOfInterest_3$protein_ID)

geneList_3 <- factor(as.integer(geneUniverse %in% genesOfInterest_3))
names(geneList_3) <- geneUniverse

myGOdata_3 <- new("topGOdata", description="time_clade3", ontology="BP", allGenes=geneList_3,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_3 <- runTest(myGOdata_3, algorithm="weight01", statistic="fisher") 

allRes_3 <- GenTable(myGOdata_3, weight = resultFisher_3,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 4

genesOfInterest_4 <- topgo_sig_4
genesOfInterest_4 <- as.character(genesOfInterest_4$protein_ID)

geneList_4 <- factor(as.integer(geneUniverse %in% genesOfInterest_4))
names(geneList_4) <- geneUniverse

myGOdata_4 <- new("topGOdata", description="time_clade4", ontology="BP", allGenes=geneList_4,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_4 <- runTest(myGOdata_4, algorithm="weight01", statistic="fisher") 

allRes_4 <- GenTable(myGOdata_4, weight = resultFisher_4,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 5

genesOfInterest_5 <- topgo_sig_5
genesOfInterest_5 <- as.character(genesOfInterest_5$protein_ID)

geneList_5 <- factor(as.integer(geneUniverse %in% genesOfInterest_5))
names(geneList_5) <- geneUniverse

myGOdata_5 <- new("topGOdata", description="time_clade5", ontology="BP", allGenes=geneList_5,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_5 <- runTest(myGOdata_5, algorithm="weight01", statistic="fisher") 


allRes_5 <- GenTable(myGOdata_5, weight = resultFisher_5,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]
sub_allRes3 <- allRes_3[which(allRes_3$Annotated >=5 & allRes_3$weight <= 0.05),]
sub_allRes4 <- allRes_4[which(allRes_4$Annotated >=5 & allRes_4$weight <= 0.05),]
sub_allRes5 <- allRes_5[which(allRes_5$Annotated >=5 & allRes_5$weight <= 0.05),]


#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2,sub_allRes3,sub_allRes4,sub_allRes5)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

sub_allRes <- merge(sub_allRes, sub_allRes3[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[5] <- "clade3"

sub_allRes <- merge(sub_allRes, sub_allRes4[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[6] <- "clade4"

sub_allRes <- merge(sub_allRes, sub_allRes5[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[7] <- "clade5"

write.csv(sub_allRes, "Time_Clades_SigGO.csv", quote = FALSE, row.names = FALSE)
```

#get GO slims for clade terms
```{r}
library(GSEABase)
#make list of unique GO terms without a protein ID column
GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]
GO_clade3 <- sub_allRes[which(!is.na(sub_allRes$clade3)),"GO.ID"]
GO_clade4 <- sub_allRes[which(!is.na(sub_allRes$clade4)),"GO.ID"]
GO_clade5 <- sub_allRes[which(!is.na(sub_allRes$clade5)),"GO.ID"]



#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)
collection_clade3 <- GOCollection(GO_clade3)
collection_clade4 <- GOCollection(GO_clade4)
collection_clade5 <- GOCollection(GO_clade5)

#I downloaded goslim_generic.obo from http://geneontology.org/docs/go-subset-guide/
#then i moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))
slims_clade3 <- data.frame(goSlim(collection_clade3, slim, "BP"))
slims_clade4 <- data.frame(goSlim(collection_clade4, slim, "BP"))
slims_clade5 <- data.frame(goSlim(collection_clade5, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2, slims_clade3, slims_clade4, slims_clade5)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

slims <- merge(slims, slims_clade3[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[4] <- "clade3"

slims <- merge(slims, slims_clade4[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[5] <- "clade4"

slims <- merge(slims, slims_clade5[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[6] <- "clade5"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]

write.csv(slims, "Time_Clades_SigGOSlim.csv",quote = F, row.names = F)

slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))




#library(viridis)
#col = rev(viridis(11))
col = c("#FFFFFFFF","#C2DF23FF","#85D54AFF","#51C56AFF","#2BB07FFF","#1E9B8AFF","#25858EFF","#2D708EFF","#38598CFF","#433E85FF","#482173FF","#440154FF")
breaks <- c(-1,0,1,2,3,4,5,6,7,8,9,10,11)

ColSideColors<- c("#000000","#00FF00","#808080","#8000FF","#80FFFF")
pdf(file="GOSlimheatmapTime.pdf", width=8, height=8)
heatmap.2(slims_m,margins = c(5,20),cexRow = 0.8, Colv=NA,revC = TRUE,breaks= breaks,col = col,density.info = "none", trace = "none", scale = "none",ColSideColors = ColSideColors)
dev.off()

```












COMMENT OUT TIME OR TEMP IN THIS CHUNK BEFORE RUNNING CODE
**Extract GO IDs for different clades**
NEED TO CHANGE FILE NAMES FOR SAVING DATA
```{r}
#time clades
silo39_clade1_GO <- STACKED_silo39_sig_pro_GOid_term[which(STACKED_silo39_sig_pro_GOid_term$protein_ID %in% silo39_time_ASCA[which(silo39_time_ASCA$ClusterID == 1),"protein_ID"]),]

write.csv(silo39_clade1_GO, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo39_time_clade1_GO.csv", quote = FALSE, row.names = FALSE)

silo39_clade2_GO <- STACKED_silo39_sig_pro_GOid_term[which(STACKED_silo39_sig_pro_GOid_term$protein_ID %in% silo39_time_ASCA[which(silo39_time_ASCA$ClusterID == 2),"protein_ID"]),]

write.csv(silo39_clade2_GO, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo39_time_clade2_GO.csv", quote = FALSE, row.names = FALSE)

silo39_clade3_GO <- STACKED_silo39_sig_pro_GOid_term[which(STACKED_silo39_sig_pro_GOid_term$protein_ID %in% silo39_time_ASCA[which(silo39_time_ASCA$ClusterID == 3),"protein_ID"]),]

write.csv(silo39_clade3_GO, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo39_time_clade3_GO.csv", quote = FALSE, row.names = FALSE)

silo39_clade4_GO <- STACKED_silo39_sig_pro_GOid_term[which(STACKED_silo39_sig_pro_GOid_term$protein_ID %in% silo39_time_ASCA[which(silo39_time_ASCA$ClusterID == 4),"protein_ID"]),]

write.csv(silo39_clade4_GO, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo39_time_clade4_GO.csv", quote = FALSE, row.names = FALSE)

silo39_clade5_GO <- STACKED_silo39_sig_pro_GOid_term[which(STACKED_silo39_sig_pro_GOid_term$protein_ID %in% silo39_time_ASCA[which(silo39_time_ASCA$ClusterID == 5),"protein_ID"]),]

write.csv(silo39_clade5_GO, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo39_time_clade5_GO.csv", quote = FALSE, row.names = FALSE)

```

**TOP GO fisher test style enrichment**
```{r}
#save total number of clade proteins and universe proteins as objects
total_clade1_prots <- length(unique(silo39_clade1_GO$protein_ID))
total_clade2_prots <- length(unique(silo39_clade2_GO$protein_ID))

#COMMENT THESE 3 OUT WHEN RUNNING TEMPERATURE
total_clade3_prots <- length(unique(silo39_clade3_GO$protein_ID))
total_clade4_prots <- length(unique(silo39_clade4_GO$protein_ID))
total_clade5_prots <- length(unique(silo39_clade5_GO$protein_ID))


total_universe_prots <- length(unique(STACKED_silo39_ALL_pro_GOid_term$protein_ID))

#define temp or time
k <- "temp"
n <- "time"

#define GO enrichment pval cutoff
pval <- 0.005

#loop through each unique GO ID and perform a fisher test (comparing representation of GO ID in clade to its representation in the universe. In other words testing whether the fraction of significant proteins annotated to the GO ID is different than the fraction of universe proteins annotated to the GO ID)
for(j in (1:5)){
  clade_data <- read.csv(paste0("silo39_",n,"_clade",j,"_GO.csv"), stringsAsFactors = FALSE)
  uniq_clade_data <- unique(clade_data$GO)

  #make table with unique GO ID and # proteins with GO ID
  GO_enrich <- data.frame(matrix(0,nrow=length(uniq_clade_data),ncol=3))
  colnames(GO_enrich) <- c("GO","p.value","p.adj")
  for(i in (1:length(uniq_clade_data))){
    num_sig_prots <- nrow(clade_data[grep(uniq_clade_data[[i]],clade_data$GO),])
    num_prots <- nrow(STACKED_silo39_ALL_pro_GOid_term[grep(uniq_clade_data[[i]],STACKED_silo39_ALL_pro_GOid_term$GO),])
    cont_table <- matrix(c(num_sig_prots,num_prots,total_clade1_prots,total_universe_prots),nrow=2,dimnames = list(c("sig prots","prots"),c("total sig prots","prot univ")))
    test <- fisher.test(cont_table) 
    GO_enrich$GO[i] <- uniq_clade_data[[i]]
    GO_enrich$p.value[i] <- test$p.value
    GO_enrich$p.adj[i] <- p.adjust(test$p.value, method = "fdr", n = length(uniq_clade_data))
  }
  GO_enrich_0.005 <- data.frame(GO_enrich[which(GO_enrich$p.value<=0.005),c("GO","p.value","p.adj")])
  print(nrow(GO_enrich_0.005))
  write.csv(GO_enrich_0.005,paste0("GO_enrich_0.005_silo39_",n,"_clade",j,".csv"),quote = FALSE, row.names = FALSE)
}

```

```{r}
GO_enrich_0.005_silo39_time_clade_1 <- read.csv(paste0("GO_enrich_0.005_silo39_time_clade",1,".csv"), stringsAsFactors = FALSE)
GO_enrich_0.005_silo39_time_clade_2 <- read.csv(paste0("GO_enrich_0.005_silo39_time_clade",2,".csv"), stringsAsFactors = FALSE)
GO_enrich_0.005_silo39_time_clade_3 <- read.csv(paste0("GO_enrich_0.005_silo39_time_clade",3,".csv"), stringsAsFactors = FALSE)
GO_enrich_0.005_silo39_time_clade_4 <- read.csv(paste0("GO_enrich_0.005_silo39_time_clade",4,".csv"), stringsAsFactors = FALSE)
GO_enrich_0.005_silo39_time_clade_5 <- read.csv(paste0("GO_enrich_0.005_silo39_time_clade",5,".csv"), stringsAsFactors = FALSE)


GO_enrich_0.005_silo39_time_clade_1$clade <- 1
GO_enrich_0.005_silo39_time_clade_2$clade <- 2
GO_enrich_0.005_silo39_time_clade_3$clade <- 3
GO_enrich_0.005_silo39_time_clade_4$clade <- 4
GO_enrich_0.005_silo39_time_clade_5$clade <- 5


STACKED_allTime <- rbind(GO_enrich_0.005_silo39_time_clade_1,GO_enrich_0.005_silo39_time_clade_2,GO_enrich_0.005_silo39_time_clade_3, GO_enrich_0.005_silo39_time_clade_4, GO_enrich_0.005_silo39_time_clade_5)

STACKED_allTime_sig <- STACKED_allTime[which(STACKED_allTime$p.adj <= 0.01),]

STACKED_allTime_sig_noFDR <- STACKED_allTime[which(STACKED_allTime$p.value <= 0.001),]

#create a list of character vectors called terms from enriched GO IDs
  terms <- as.list(STACKED_allTime_sig_noFDR$GO)
  #Get GO info in ontologyX database (GO.obo) for terms in "terms" list
  GO_IC_data <- get_term_info_content(go, term_sets = as.vector(STACKED_allTime_sig_noFDR$GO))
  #Subset terms to match 
  terms <- terms[terms %in% names(GO_IC_data)]

  
STACKED_allTime_sig_noFDR <- STACKED_allTime_sig_noFDR[which(STACKED_allTime_sig_noFDR$GO %in% terms),] 

STACKED_allTime_sig_noFDR_allpadj <- STACKED_allTime[which(STACKED_allTime$GO %in% terms),]

#Get the full GO term info for each GO ID
for(i in 1:nrow(STACKED_allTime_sig_noFDR_allpadj)){
  STACKED_allTime_sig_noFDR_allpadj$term[i] <- go$name[[STACKED_allTime_sig_noFDR_allpadj$GO[i]]]
}

#remove commas from GO terms so it wont interfere with saving data as .csv
STACKED_allTime_sig_noFDR_allpadj$term <- gsub(",",";", STACKED_allTime_sig_noFDR_allpadj$term)


#remove duplicates
allTime_sig_noFDR <- tidyr::spread(STACKED_allTime_sig_noFDR_allpadj,"clade", "p.adj")
allTime_sig_noFDR <- allTime_sig_noFDR[which(!duplicated(allTime_sig_noFDR$term)),]

allTime_sig_noFDR_m <- allTime_sig_noFDR[,4:8]
rownames(allTime_sig_noFDR_m) <- allTime_sig_noFDR[,"term"]

allTime_sig_noFDR_m <- data.matrix(allTime_sig_noFDR_m)

heatmap.2(allTime_sig_noFDR_m,cexRow = 0.25, cexCol = 0.8, Colv=NA, col = heat.colors, na.color = "white", density.info = "none", trace = "none", scale = "row")


heatmap3(data.matrix(allTime_sig_noFDR_m))

allTime_sig_noFDR[,4:8] <- as.numeric(allTime_sig_noFDR[,4:8])

```

GSEA to reduce terms
```{r}
library(GSEABase)
#BP first
myCollection.clade1 <- GOCollection(STACKED_allTime_sig_noFDR[which(STACKED_allTime_sig_noFDR$clade ==1),"GO"])

myCollection.clade2 <- GOCollection(STACKED_allTime_sig_noFDR[which(STACKED_allTime_sig_noFDR$clade ==2),"GO"])

myCollection.clade3 <- GOCollection(STACKED_allTime_sig_noFDR[which(STACKED_allTime_sig_noFDR$clade ==3),"GO"])

myCollection.clade4 <- GOCollection(STACKED_allTime_sig_noFDR[which(STACKED_allTime_sig_noFDR$clade ==4),"GO"])

myCollection.clade5 <- GOCollection(STACKED_allTime_sig_noFDR[which(STACKED_allTime_sig_noFDR$clade ==5),"GO"])



slims_clade1 <- data.frame(goSlim(myCollection.clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(myCollection.clade2, slim, "BP"))
slims_clade3 <- data.frame(goSlim(myCollection.clade3, slim, "BP"))
slims_clade4 <- data.frame(goSlim(myCollection.clade4, slim, "BP"))
slims_clade5 <- data.frame(goSlim(myCollection.clade5, slim, "BP"))

#add clade info to slims
colnames(slims_clade1) <- c("1","percent","Term")
colnames(slims_clade2) <- c("2","percent","Term")
colnames(slims_clade3) <- c("3","percent","Term")
colnames(slims_clade4) <- c("4","percent","Term")
colnames(slims_clade5) <- c("5","percent","Term")



#bind all slim columns together
slims_all_clades <- merge(slims_clade1[,-2], slims_clade2[,-2], by = "Term", all = TRUE)

slims_all_clades <- merge(slims_all_clades, slims_clade3[,-2], by = "Term", all = TRUE)

slims_all_clades <- merge(slims_all_clades, slims_clade4[,-2], by = "Term", all = TRUE)

slims_all_clades <- merge(slims_all_clades, slims_clade5[,-2], by = "Term", all = TRUE)

#remove rows with 0 in all columns
rownames(slims_all_clades) <- slims_all_clades$Term
slims_all_clades <- slims_all_clades[,2:6]

slims_all_clades <- slims_all_clades[which(apply(slims_all_clades,1,sum)!= 0),]
#remove biological process term
slims_all_clades <- slims_all_clades[which(rownames(slims_all_clades)!= "biological_process"),]

#remove rows with row sum = 1

slims_all_clades <- slims_all_clades[which(apply(slims_all_clades,1,sum)!= 1),]


heatmap3(data.matrix(slims_all_clades), Colv = NA, cexRow = 0.8)

library(RColorBrewer)
heatmap.2(data.matrix(slims_all_clades),cexRow = 0.75, Colv=NA, col = brewer.pal(5,"YlOrRd") , na.color = "white",density.info = "none", trace = "none")

heatmap.2(data.matrix(slims_all_clades),cexRow = 0.75, Colv=NA, col = brewer.pal(5,"YlOrRd") , na.color = "white",density.info = "none", trace = "none")

slims_all_clades[slims_all_clades == 0] <- NA

heatmap.2(data.matrix(slims_all_clades),cexRow = 0.75, Colv=NA, col = my_palette,density.info = "none", trace = "none")



my_palette <- colorRampPalette(c("white", "pink", "red"))(n = 11)
col= colorRampPalette(c("white", "red"))(12)

```
**GO SEMSIM OF ENRICHED GENES**
```{r}
#load GO data and GO info from OntologyIndex package
data(go)
data(GO_IC)

for(j in (1:5)){
  GO_enrich_0.05 <- read.csv(paste0("GO_enrich_0.005_silo39_time_clade",j,".csv"), stringsAsFactors = FALSE)
 #create a list of character vectors called terms from enriched GO IDs
  terms <- as.list(GO_enrich_0.05$GO)
  #Get GO info in ontologyX database (GO.obo) for terms in "terms" list
  GO_IC_data <- get_term_info_content(go, term_sets = as.vector(GO_enrich_0.05$GO))
  #Subset terms to match 
  terms <- terms[terms %in% names(GO_IC_data)]
  #create a semantic similarity matrix with enriched GO IDs
  sim_matrix <- get_sim_grid(ontology=go,information_content=GO_IC_data,term_sets=terms)
  #add row and column names that are the GO IDs
  rownames(sim_matrix) <- unlist(terms)
  colnames(sim_matrix) <- unlist(terms)
  
  #save sim matrix
  write.csv(sim_matrix,paste0("enrichGOsemsim_matrix_","n_",j,".csv"), quote = FALSE)
  
  #plot heat map of simantic similarities
  pdf(file=paste0("GOsimsem_heatmap_silo39_time_clade",j,".pdf"), width=10, height=8)
  heatmap3(sim_matrix, scale = "none")
  dev.off()
}


```

**Finding clusters in heatmap**
Dendrogram should be inspected manually with heatmap to determine where the tree should be cut to get the appropriate number of clusters
```{r}
#find clusters in heatmap

#define clade
j=5
#read in matrix data
sim_matrix <- read.csv(paste0("enrichGOsemsim_matrix_","n_",j,".csv"), stringsAsFactors = FALSE)

#format matrix data for as.dist; data.matrix throws an error because the row names get removed; they are added back in 4 lines below
sim_matrix <- data.matrix(sim_matrix)
sim_matrix <- sim_matrix[,-1]
colnames(sim_matrix) <- gsub("\\.",":", colnames(sim_matrix))
rownames(sim_matrix) <- colnames(sim_matrix)
#plot dendrogram  
hm <- as.dist(1-cor(t(sim_matrix), use="pa"))
hm2 <- hclust(hm, method = 'complete')
plot(hm2)

#cut <- 0.29 for time clades 1-3;0.24 for clade 4; 0.35 for clade 5
cut <- 0.35
abline(h=cut, col = "red")

############################################################
#make a list of proteins with cluster IDs based on where the tree is cut

#for complete clustering
mycl <- cutree(hm2,h=cut)

#assign colors to cluster IDs
clusterCols <- colorRamps::primary.colors(length(unique(mycl)))
myClusterSideBar <- clusterCols[mycl]

#replot and save heatmap with clades
pdf(file=paste0("GOsimsem_heatmap_silo39_",n,"_clade",j,".pdf"), width=10, height=8)
heatmap3(sim_matrix,cexRow = 0.5, cexCol = 0.5, RowSideColors=myClusterSideBar,RowAxisColors=1,hclustfun=hclust, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), method = "complete", scale = "none")
dev.off()

#get color name
clusterColor <- lapply(myClusterSideBar, color.id)
clusterColor <- lapply(clusterColor, `[[`,1)
clusterColor <- data.frame(unlist(clusterColor), stringsAsFactors = FALSE)

#merge cluster color with GO ID and cluster number
foo <- cbind(data.frame(mycl), clusterColor)
foo$GO <- rownames(foo)
colnames(foo) <- c("Clust_Num","Clust_Color","GO")

#set values < 0.5 or 1 in sim_matrix to 0, convert to dataframe and add GO IDs as column
sim_matrix1 <- sim_matrix
sim_matrix1[sim_matrix1<0.5] <- 0
sim_matrix1[sim_matrix1==1] <- 0

sim_matrix1 <- data.frame(sim_matrix1)
sim_matrix1$GO <- row.names(sim_matrix1)

#merge clade info with sim_matrix
sim_matrix_clade <- merge(foo,data.frame(sim_matrix1), by = "GO")

#sum the semantic similarity scores for each GO ID
sim_matrix_clade$sumsim <- apply(sim_matrix_clade[,5:ncol(sim_matrix_clade)],1,sum)
#exclude any terms that summed to 0; this happens because many terms are set to 0 in the step above and remove individual GO columns now that we have the sum for each term
sim_matrix_clade <- sim_matrix_clade[which(sim_matrix_clade$sumsim!=0),c(1:3,ncol(sim_matrix_clade))]

#make a data frame with the cluster color and the semantic similarity sum of all term sums
color_sums <- aggregate(sim_matrix_clade$sumsim, by=list(Category=sim_matrix_clade$Clust_Color), FUN=sum)
#rename columns
colnames(color_sums) <- c("Clust_Color","semsimsum")
#add color sum to the abbreviated sim matrix
sim_matrix_clade_sums <- merge(sim_matrix_clade,color_sums, by = "Clust_Color")
#Get the full GO term info for each GO ID
for(i in 1:nrow(sim_matrix_clade_sums)){
  sim_matrix_clade_sums$term[i] <- go$name[[sim_matrix_clade_sums$GO[i]]]
}

#remove commas from GO terms so it wont interfere with saving data as .csv
sim_matrix_clade_sums$term <- gsub(",",";", sim_matrix_clade_sums$term)

#read in GO slim info to add to sim_matrix_clade_sums data frame
srlabslim <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/background/GOSlim_terms.csv", stringsAsFactors = FALSE)
colnames(srlabslim)[1] <- "GO"

#some GO terms have muliple GO slims so for these, combine GO slims into one line and separate them with a ;
agg_df <- merge(sim_matrix_clade_sums,srlabslim[,c("GO","GOSlim_bin")],by = "GO",all.x = TRUE)
agg_df <- aggregate(GOSlim_bin~GO, agg_df, paste, collapse = ";")
#merge GOslim data with sim_matrix_clade_sums data frame
sim_matrix_clade_sums <-merge(sim_matrix_clade_sums,agg_df, by = "GO", all.x = TRUE)
#add p.value info to data frame
GO_enrich_0.05 <- read.csv(paste0("GO_enrich_0.005_silo39_",n,"_clade",j,".csv"), stringsAsFactors = FALSE)
sim_matrix_clade_sums <- merge(sim_matrix_clade_sums,GO_enrich_0.05, by = "GO", all.x = TRUE)
#order data by semantic similarity sum
sim_matrix_clade_sums <- sim_matrix_clade_sums[order(sim_matrix_clade_sums$semsimsum,decreasing = TRUE),]


write.csv(sim_matrix_clade_sums,paste0("silo39_",n,"_clade",j,"_GOenrich_clust.csv"), quote = FALSE, row.names = FALSE)

###GO terms and pvalues are then uploaded into revigo with 0.5 (small setting)

```