---
title: "ASCA on average NSAF values of filtered proteins"
author: "Shelly Trigg"
date: "1/16/2019"
output: rmarkdown::github_document
---

load libraries
```{r}
library(dplyr)
library(tidyr)
library(MetStaT)
library(ggplot2)
library(heatmap3)
```

load data
```{r}
#NSAF data from filtered proteins
data <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/silo3and9_nozerovals_noincnstprot.csv", stringsAsFactors = FALSE)
```

load data with annotations so protein IDs can be substituted
This chunk makes a list of protein IDs and unique shortened Entry Names
```{r}
#data with uniprot annotations
data_w_uniprot <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/gigaton-uniport-table_GOterms-sperated.csv", stringsAsFactors = FALSE)
#subset only Protein IDs and Entry names
data_w_uniprot <- data_w_uniprot[,c("Protein.ID", "Entry.name")]
#replace pipe character with nothing; remember \\ to escape the | character
data_w_uniprot$Protein.ID <- gsub("\\|",".", data_w_uniprot$Protein.ID)
#replace all characters following '_' with nothing in the entry name column
data_w_uniprot$Entry.name.simple <- gsub("\\_.*","", data_w_uniprot$Entry.name)
#replace 'none' with 'unknown'
data_w_uniprot$Entry.name.simple <- gsub("None","Unknown", data_w_uniprot$Entry.name.simple)

#there are duplicated Entry names for unique protein IDs, so we need to make unique Entry names by adding .1, .2 etc.
#https://stackoverflow.com/questions/16646446/renaming-duplicate-strings-in-r

data_w_uniprot$Entry.name.new <- make.names(data_w_uniprot$Entry.name.simple, unique = TRUE)
```

subset the data for only CHOYP proteins; this removes contaminant proteins
```{r}
#make a dataframe with just CHOYP proteins
choyp_data <- data[,c(1:2,grep("CHOYP", colnames(data)))]
#make a seperate data frame with contaminant proteins
contam_data <- data[,-grep("CHOYP", colnames(data))]
```

perform ASCA
```{r}
#create matrix to pass to ASCA command, excluding the silo and time info
ASCA_X <- as.matrix(choyp_data[,-c(1:2)])
#create matrix to pass to ASCA command with only the silo and time info
ASCA_F <- as.matrix(choyp_data[,c(1:2)])
#perform ASCA
ASCA <- ASCA.Calculate(ASCA_X, ASCA_F, equation.elements = "1,2,12", scaling = FALSE)
```

plot PCAs from ASCA
```{r}
#plot PCA for factor 1, which is time in this case
ASCA.PlotScoresPerLevel(ASCA, ee = "1", pcs = "1,2")
#plot PCA for factor 2, which is temperature in this case
ASCA.PlotScoresPerLevel(ASCA, ee = "2", pcs = "1,2")
#print the ASCA summary
ASCA.GetSummary(ASCA)
```

Analysis of proteins affected by temperature
```{r}
#extract protein names from ASCA data; these will be combined with loadings values; the order is maintained
protnames <- data.frame(colnames(ASCA$data))
#combine protein names with ASCA loadings for PC2 for temperature, since this component showed the greatest separation between temperatures
d <- cbind(protnames, ASCA$`2`$svd$v[,2])
#rename the columns
colnames(d) <- c("protein", "PC2loadings")
#make a dataframe of proteins with PC loadings greater than zero for the loadings plot
d_great <- d[which(d$PC2loadings > 0),]
#make a dataframe of proteins with PC loadings less than zero for the loadings plot
d_less <- d[which(d$PC2loadings < 0),]
#plot PC2 loadings
plot(d_great[order(d_great$PC2loadings, decreasing = TRUE),2], xlab = "protein", ylab = "PC2 Loadings")
plot(d_less[order(d_less$PC2loadings, decreasing = TRUE),2],xlab = "protein", ylab = "PC2 Loadings")
```

To pull out proteins affected by temperature based on their influence in seperating treatment groups on PC2 of the temperature PCA, I picked an absolute value loadings threshold of 0.025. This means any protein that had a loadings value > 0.025 or < -0.025 was selected.
```{r}
#make list of proteins with temperature PC1 loadings values >= 0.04
cutd <- d[which(abs(d$PC2loadings) >= 0.025),]

#make a list of cutoff proteins with normalized abundance
cut_data <-data[,which(colnames(data) %in% cutd$protein)]
rownames(cut_data) <- paste(data$day, data$temp, sep="_")
```

Plot a heatmap of protein abundances (NSAF values) that are influenced by temperature having PC2 loadings values > 0.025 or < -0.025. 
```{r,fig.height=10, fig.width=10}
#make a heatmap to see if proteins cluster based on abundance over time
heatmap3(t.data.frame(cut_data),Colv = NA,cexRow = 0.4, cexCol = 0.7)
```
This heatmap shows common time points adjacent to each other.  

Make a new heatmap ordered by sequencial time points for each temperature.
```{r,fig.height=10, fig.width=10}
#diff order
cut_data_ord <- data[,c(1:3,which(colnames(data) %in% cutd$protein))]
cut_data_ord <- cut_data_ord[order(cut_data_ord$temp,cut_data_ord$day),]
rownames(cut_data_ord) <- paste(cut_data_ord$day, cut_data_ord$temp, sep="_")
cut_data_ord <- cut_data_ord[,-c(1:3)]
heatmap3(t.data.frame(cut_data_ord),Colv = NA,cexRow = 0.4, cexCol = 0.7)
```

Make a new heatmap with simpler protein names
```{r,fig.height=10, fig.width=10}
#replace CHOYP names with "entry names"
#transposed to eventually get row names as columns to match entry names to. 
#t.data.frame returns a matrix so I had convert back to data frame for all downstream rearranging because data frames are much easier to work with
cut_data_ord_t <- data.frame(t.data.frame(cut_data_ord))
#make row names which contain the protein IDs a column
cut_data_ord_t <- cbind(rownames(cut_data_ord_t), cut_data_ord_t)
#add the column name
colnames(cut_data_ord_t)[1] <- "Protein.ID"
#convert the Protein ID column to character
cut_data_ord_t[,1] <- as.character(cut_data_ord_t[,1])
#add the Entry names
cut_data_ord_t <- merge(data_w_uniprot[,c("Protein.ID","Entry.name.new")],cut_data_ord_t, by = "Protein.ID")
#make the entry names the row names
rownames(cut_data_ord_t) <- cut_data_ord_t[,"Entry.name.new"]
#remove the protein ID and entry name columns since the row names are added
cut_data_ord_t <- cut_data_ord_t[,-c(1:2)]
#this step removes the x from the column names while preserving the order of colnames
colnames(cut_data_ord_t) <- rownames(cut_data_ord)
#plot the heat map
heatmap3(as.matrix(cut_data_ord_t), Colv = NA, cexRow = 0.3, cexCol = 0.7)
```

```{r, echo = FALSE,eval=FALSE}
#Analysis of proteins affected by time (development associated proteins)

#make dataframes of factor 1 (time) PC1 loadings
d1 <- cbind(protnames, ASCA$`1`$svd$v[,1])
colnames(d1) <- c("protein", "PC1loadings")
d1_great <- d1[which(d1$PC1loadings > 0),]
d1_less <- d1[which(d1$PC1loadings < 0),]

plot(d1_great[order(d1_great$PC1loadings, decreasing = TRUE),2])
plot(d1_less[order(d1_less$PC1loadings, decreasing = TRUE),2])


#make list of proteins with temperature PC1 loadings values >= 0.04
cutd <- d[which(abs(d$PC2loadings) >= 0.025),]

#make a list of cutoff proteins with normalized abundance
cut_data <-data[,which(colnames(data) %in% cutd$protein)]
rownames(cut_data) <- paste(data$day, data$temp, sep="_")

#make a heatmap to see if proteins cluster based on abundance over time
heatmap3(as.matrix(t.data.frame(cut_data)), Colv = NA, cexRow = 0.4, cexCol = 0.7)
```

```{r, echo = FALSE, eval = FALSE}

#playing around with using derivatives to pick a loadings threshold


#https://stackoverflow.com/questions/41518870/finding-the-elbow-knee-in-a-curve
get.elbow.points.indices <- function(x, y, threshold) {
  d1 <- diff(y) / diff(x) # first derivative
  d2 <- diff(d1) / diff(x[-1]) # second derivative
  indices <- which(abs(d2) > threshold)  
  return(indices)
}



x = seq(1:nrow(d_great))
#convert this to negative sort greatest to least; this allows us to pick the first derivative
#y = d_great[order(d_great$PC2loadings, decreasing = FALSE),2] * -1
y = d_great[order(d_great$PC2loadings, decreasing = TRUE),2]

indices <- get.elbow.points.indices(x, y,0.001) # threshold for huge jump = 0.001
x[indices]
#[1]  1  2  3  4  5  6  7  8 11 12 13 14 15 16 17 18 25 26 28 29 37
y[indices]
#[1] 0.37187461 0.27551707 0.25225684 0.18367869 0.18172208 0.10977722 0.08680822 0.07276701 0.06496925
#[10] 0.06271090 0.06177204 0.05975629 0.05667319 0.05563780 0.05182859 0.05154770 0.04306674 0.04262415
#[19] 0.03987387 0.03900935 0.03375264
plot(x, y, pch=19)
points(x[indices[21]], y[indices[21]], pch=19, col='red')


x = seq(1:nrow(d_less))
y = d_less[order(d_less$PC2loadings, decreasing = TRUE),2]

indices <- get.elbow.points.indices(x, y,0.001) # threshold for huge jump = 0.001
x[indices]
#[1] 2407 2408 2418 2421 2422 2429 2431 2433 2435 2440 2441 2442 2443 2444 2445 2446 2447 2448 2449 2450 2451
#[22] 2452 2453
y[indices]
#[1] -0.02592041 -0.02598057 -0.03015541 -0.03170286 -0.03234933 -0.03757466 -0.03944236 -0.04092336
#[9] -0.04456572 -0.04964308 -0.04990225 -0.05389416 -0.05445493 -0.05711436 -0.06314595 -0.06791729
#[17] -0.07024664 -0.07398529 -0.09805340 -0.09988452 -0.10409290 -0.12862659 -0.14560657
plot(x, y, pch=19)
points(x[indices[1]], y[indices[1]], pch=19, col='red')


```

