---
title: "Clustering technical replicates"
author: "Shelly Trigg"
date: "1/11/2019"
output: rmarkdown::github_document
---

Load packages
```{r}
library(vegan)
library(ggplot2)
library(dplyr)
```

Load Abacus data, parse out ADJNSAF values, and simplify column names to just sample number
```{r}
#upload data file
ABACUSdata <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/ABACUS_output021417.tsv", sep = "\t", header=TRUE, stringsAsFactors = FALSE)
#select only columns containing ADJNSAF and Protein ID
ABACUSdata <- ABACUSdata[,c(1,grep("ADJNSAF", colnames(ABACUSdata)))]

## change column names in ABACUSdata to just sampleID
colnames(ABACUSdata) <- gsub(pattern = "X20161205_SAMPLE_", "", colnames(ABACUSdata))
colnames(ABACUSdata) <- gsub(pattern = "_ADJNSAF", "", colnames(ABACUSdata))
```

Load meta data file with temperature and day information
```{r}
#upload meta data; this was a csv file I create from Rhonda's notebook entry: https://github.com/Ellior2/Ellior2.github.io/blob/master/_posts/2017-3-11-NMDS-analysis.md
meta_data <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/Rhonda_new_sample_names.csv", header = TRUE, stringsAsFactors = FALSE)
meta_data$silo <- substr(meta_data$Contents,5,5)
meta_data$day <- substr(meta_data$SampleName,5,6)
meta_data$SampleName <- gsub(pattern = "H","",meta_data$SampleName)
meta_data$SampleName <- gsub(pattern = "C","",meta_data$SampleName)
#create a temperature column
meta_data$temp <- "temp"
for(i in 1:nrow(meta_data)){
  if(meta_data$silo[i] == "2"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "3"){
    meta_data$temp[i] <- "23"
  }
  if(meta_data$silo[i] == "9"){
    meta_data$temp[i] <- "29"
  }
  if(meta_data$silo[i] == "e"){
    meta_data$temp[i] <- "16"
  }
}
```

Reformat Abacus data for NMDS
```{r}
#Transpose- switch rows and columns
tABACUSdata <- t.data.frame(ABACUSdata[,-1])
colnames(tABACUSdata) <- ABACUSdata[,1]
tABACUSdata <- cbind(data.frame(rownames(tABACUSdata)),tABACUSdata)
colnames(tABACUSdata)[1] <- "SampleID"

#add meta data to abacus data
tABACUSdata <- merge(meta_data[,c(1,2,7,8)],tABACUSdata, by = "SampleID")

#Remove Silo 2 and day 15
silo3and9 <- tABACUSdata[which(substr(tABACUSdata$SampleName,1,2) != "S2" & tABACUSdata$day != "15"),]
#make rownames from Sample ID column so that the NMDS knows what's what
rownames(silo3and9) <- silo3and9$SampleID
#order the data frame by day and temperature so coloring the points on the plot is easier
silo3and9 <- silo3and9[order(as.numeric(silo3and9$day),silo3and9$temp),]
```

Determine if any proteins have zero ADJNSAF vals for all samples; this would be because they were in Silo 2, but not in Silo 3 or 9
```{r}
no_val_proteins <- silo3and9[,which(apply(silo3and9, 2, var) == 0)]
ncol(no_val_proteins)
```

Remove proteins if they have a zero value in all samples
```{r}
silo3and9_nozerovar <- silo3and9[,-c(1:4,which(colnames(silo3and9) %in% colnames(no_val_proteins)))]
#check to make sure it worked
ncol(silo3and9)-ncol(silo3and9_nozerovar)
```

For proteins with a zero value in any sample, replace with very small value
```{r}
silo3and9_nozerovar[silo3and9_nozerovar == 0.0000] <- 0.1000
```

try PCA

```{r}
pca <- prcomp(silo3and9_nozerovar, center = T, scale = T)
pca_meta <- cbind(silo3and9$day, silo3and9$temp, data.frame(paste(silo3and9$day,silo3and9$temp, sep = "_")),pca$x)
colnames(pca_meta)[1:3] <- c("day","temp","SampleName")
ggplot(pca_meta, aes(PC1, PC2)) + geom_point(aes(col = day, shape = temp)) + theme_bw() + ggtitle("PCA of ADJNSAF values where zeros were replaced with 0.1")
```

try PCA on log transformed values
```{r}
silo3and9_log <- log(silo3and9_nozerovar,2)
pca_log <- prcomp(silo3and9_log, center = F, scale = F)
pca_log_meta <- cbind(silo3and9$day, silo3and9$temp, data.frame(paste(silo3and9$day,silo3and9$temp, sep = "_")),pca_log$x)
colnames(pca_log_meta)[1:3] <- c("day","temp","SampleName")
ggplot(pca_log_meta, aes(PC1, PC2)) + geom_point(aes(col = day, shape = temp)) + theme_bw() + ggtitle("PCA of log ADJNSAF values with zeros replaced with 0.1")
```

Make MDS dissimilarity matrix
```{r}
nmds.silo3and9 <- metaMDS(silo3and9_nozerovar, distance = 'euclidean', k = 2, trymax = 3000, autotransform = FALSE)

#make data frame of NMDS scores
nmds.silo3and9.scores <- cbind(silo3and9$day, silo3and9$temp,data.frame(scores(nmds.silo3and9)))
colnames(nmds.silo3and9.scores)[1:2] <- c("day","temp")
ggplot(nmds.silo3and9.scores, aes(NMDS1, NMDS2)) + geom_point(aes(col = day, shape = temp)) + theme_bw() + ggtitle("NMDS of ADJNSAF values with zeros replaced with 0.1")
```

Make MDS dissimilarity matrix with log tranformed ADJNSAF values
```{r}
nmds.silo3and9_log <- metaMDS(silo3and9_log, distance = 'euclidean', k = 2, trymax = 3000, autotransform = FALSE)

#make data frame of NMDS scores
nmds.silo3and9_log.scores <- cbind(silo3and9$day, silo3and9$temp,data.frame(scores(nmds.silo3and9_log)))
colnames(nmds.silo3and9_log.scores)[1:2] <- c("day","temp")
ggplot(nmds.silo3and9_log.scores, aes(NMDS1, NMDS2)) + geom_point(aes(col = day, shape = temp)) + theme_bw() + ggtitle("NMDS of log ADJNSAF values with zeros replaced with 0.1")
```

Make MDS dissimilarity matrix with log transformed ADJNSAF values and bray curtis distance
```{r}
nmds.silo3and9_log_bray <- metaMDS(silo3and9_log, distance = 'bray', k = 2, trymax = 3000, autotransform = FALSE)

#make data frame of NMDS scores
nmds.silo3and9_log_bray.scores <- cbind(silo3and9$day, silo3and9$temp,data.frame(scores(nmds.silo3and9_log_bray)))
colnames(nmds.silo3and9_log_bray.scores)[1:2] <- c("day","temp")
ggplot(nmds.silo3and9_log_bray.scores, aes(NMDS1, NMDS2)) + geom_point(aes(col = day, shape = temp)) + theme_bw() + ggtitle("bray curtis NMDS of log ADJNSAF values with zeros replaced with 0.1")

