---
title: "DaysxTime_normalization"
author: "Shelly Trigg"
date: "5/23/2019"
output: rmarkdown::github_document
---

load libraries
```{r}
library(tidyr)
library(ggplot2)
library(MetStaT)
library(heatmap3)
```

```{r}
NSAF <- read.csv("~/Documents/GitHub/paper-OysterSeed-TimeXTemp/Data/silo3and9_nozerovals_NSAF_AVGs.csv", stringsAsFactors = FALSE)
NSAF <- NSAF[,c(1,2,grep("CHOYP",colnames(NSAF)))]
```

Create tempXdays dataframe and merge it with NSAF data
```{r}
tempXdays <- data.frame(NSAF$day * NSAF$temp)

NSAF <- cbind(tempXdays, NSAF)
colnames(NSAF)[1] <- "tempXdays"
```

normalize all NSAFs by tempXdays
```{r}
NSAF_normTxD <- data.frame()

for(i in 1:nrow(NSAF)){
row <- NSAF[i,4:ncol(NSAF)]/(NSAF$tempXdays[i])
NSAF_normTxD <- rbind(NSAF_normTxD, row)
}
```

add temp and day info back
```{r}
NSAF_normTxD <- cbind(NSAF[,2:3], NSAF_normTxD)
```


**Perform ASCA**
```{r}
#create matrix to pass to ASCA command, excluding the silo and time info
ASCA_X <- as.matrix(NSAF_normTxD[,-c(1:2)])
#create matrix to pass to ASCA command with only the silo and time info
ASCA_F <- as.matrix(NSAF_normTxD[,c(1:2)])
#perform ASCA
ASCA <- ASCA.Calculate(ASCA_X, ASCA_F, equation.elements = "1,2,12", scaling = FALSE)
```

Here is a summary of the ASCA results (e.g. variance explained by different factors; factor 1= time (days), factor 2 = temperature, interaction = interaction of time and temperature)
```{r}
#print the ASCA summary
ASCA.GetSummary(ASCA)
```

### Plot PCAs from ASCA  

**This first plot is the time (days) effect PCA**
```{r avgNSAF_PCA_timeEffect_plot}
#plot PCA for factor 1, which is time in this case
ASCA.PlotScoresPerLevel(ASCA, ee = "1", pcs = "1,2")
```

**This next plot is the temperature effect PCA**
```{r avgNSAF_PCA_tempEffect_plot}
#plot PCA for factor 2, which is temperature in this case
ASCA.PlotScoresPerLevel(ASCA, ee = "2", pcs = "1,2")
```

**This next plot is the time x temp interaction effect PCA**
```{r avgNSAF_PCA_timeXtempEffect_plot}
#plot PCA for factor interaction, which is time x temp in this case
timextemp_PC12 <- data.frame(ASCA$`12`$svd$t[,c(1,2)])
timextemp_PC12 <- cbind(data.frame(ASCA$`12`$level.combinations$row.patterns), timextemp_PC12)
colnames(timextemp_PC12)<- c("day","temp","PC1","PC2")
timextemp_PC12$day <- as.character(timextemp_PC12$day)
timextemp_PC12$temp <- as.character(timextemp_PC12$temp)
ggplot(timextemp_PC12, aes(PC1, PC2)) + geom_point(aes(col = day, shape = temp, size = 3)) + theme_bw() + ggtitle("PC1 vs PC2 for time x temperature interaction effect") + theme(plot.title = element_text(face = "bold")) + xlab(paste("PC1"," (",formatC(ASCA$`12`$svd$var.explained[1] * 100,digits=2,format="f"),"%)", sep = "")) + ylab(paste("PC2"," (",formatC(ASCA$`12`$svd$var.explained[2] * 100,digits=2,format="f"),"%)", sep = ""))
```

### Analysis of proteins affected by temperature

Because the temperature effect PCA show the most separation between 23C and 29C in PC2, we will look at those loadings.

**PC1 loadings for temperature effect**
```{r avgNSAF_PCA_tempEffect_PC2loadings, echo = FALSE}
#extract protein names from ASCA data; these will be combined with loadings values; the order is maintained
protnames <- data.frame(colnames(ASCA$data))
#combine protein names with ASCA loadings for PC2 for temperature, since this component showed the greatest separation between temperatures
d <- cbind(protnames, ASCA$`2`$svd$v[,1])
#rename the columns
colnames(d) <- c("protein", "PC1loadings")
#make a dataframe of proteins with PC loadings greater than zero for the loadings plot
d_great <- d[which(d$PC1loadings > 0),]
#make a dataframe of proteins with PC loadings less than zero for the loadings plot
d_less <- d[which(d$PC1loadings < 0),]
#write out PC1 loadings for ASCA$2$svd$[,1]
#write.csv(d, "~/Documents/GitHub/paper-OysterSeed-TimeXTemp/Data/Supplementary_Data6.csv", row.names = FALSE, quote = FALSE)
#plot PC1 loadings
plot(d_great[order(d_great$PC1loadings, decreasing = TRUE),2], xlab = "protein", ylab = "PC1 Loadings")
abline(h=0.01, col = "red")
plot(d_less[order(d_less$PC1loadings, decreasing = TRUE),2],xlab = "protein", ylab = "PC1 Loadings")
abline(h=-0.035, col = "red")

```

To pull out proteins affected by temperature based on their influence in seperating treatment groups on PC2 of the temperature PCA, I picked an absolute value loadings threshold of 0.025. This means any protein that had a loadings value > 0.025 or < -0.025 was selected.
```{r, echo = FALSE}
#make list of proteins with temperature PC1 loadings values >= 0.025
cutd <- d[which(d$PC1loadings >= 0.01 | d$PC1loadings <= -0.035),]
#make a list of cutoff proteins with normalized abundance
cut_data <-NSAF_normTxD[,which(colnames(NSAF_normTxD) %in% cutd$protein)]
rownames(cut_data) <- paste(NSAF_normTxD$day, NSAF_normTxD$temp, sep="_")
```


heatmap3(t(cut_data))
heatmap3(t(cut_data), Colv = NA)

**PC1 loadings for temperature x time effect**
```{r avgNSAF_PCA_tempxtimeEffect_PC2loadings, echo = FALSE}
#extract protein names from ASCA data; these will be combined with loadings values; the order is maintained
protnames <- data.frame(colnames(ASCA$data))
#combine protein names with ASCA loadings for PC2 for temperature, since this component showed the greatest separation between temperatures
d <- cbind(protnames, ASCA$`12`$svd$v[,1])
#rename the columns
colnames(d) <- c("protein", "PC1loadings")
#make a dataframe of proteins with PC loadings greater than zero for the loadings plot
d_great <- d[which(d$PC1loadings > 0),]
#make a dataframe of proteins with PC loadings less than zero for the loadings plot
d_less <- d[which(d$PC1loadings < 0),]
#write out PC1 loadings for ASCA$2$svd$[,1]
#write.csv(d, "~/Documents/GitHub/paper-OysterSeed-TimeXTemp/Data/Supplementary_Data6.csv", row.names = FALSE, quote = FALSE)
#plot PC1 loadings
plot(d_great[order(d_great$PC1loadings, decreasing = TRUE),2], xlab = "protein", ylab = "PC1 Loadings")
abline(h=0.05, col = "red")
plot(d_less[order(d_less$PC1loadings, decreasing = TRUE),2],xlab = "protein", ylab = "PC1 Loadings")
abline(h=-0.035, col = "red")

```

To pull out proteins affected by temperature based on their influence in seperating treatment groups on PC2 of the temperature PCA, I picked an absolute value loadings threshold of 0.025. This means any protein that had a loadings value > 0.025 or < -0.025 was selected.
```{r, echo = FALSE}
#make list of proteins with temperature PC1 loadings values >= 0.025
cutd <- d[which(d$PC1loadings >= 0.05 | d$PC1loadings <= -0.035),]
#make a list of cutoff proteins with normalized abundance
cut_data12 <-NSAF_normTxD[,which(colnames(NSAF_normTxD) %in% cutd$protein)]
rownames(cut_data12) <- paste(NSAF_normTxD$day, NSAF_normTxD$temp, sep="_")
```

plot a  heatmap of proteins with high PC1 loadings values in the ASCA time-temp interaction effect plot
```{r}
heatmap3(t(cut_data12), Colv = NA, cexRow = 0.5)
```

pca
```{r}
pca <- prcomp(log(NSAF_normTxD[,-c(1:2)],2))
pca_meta <- cbind(NSAF_normTxD[,c(1:2)],pca$x)
pca_meta$day <- factor(pca_meta$day)
pca_meta$temp <- factor(pca_meta$temp)

ggplot(pca_meta, aes(PC1, PC2)) + geom_point(aes(col = day, shape = temp)) + theme_bw() + ggtitle("PCA of ADJNSAF values scaled by day*time")
```



PCA with log scaled NSAFs
```{r}
pca <- prcomp(log(NSAF_normTxD[,-c(1:2)],2))
pca_meta <- cbind(NSAF_normTxD[,c(1:2)],pca$x)
pca_meta$day <- factor(pca_meta$day)
pca_meta$temp <- factor(pca_meta$temp)

ggplot(pca_meta, aes(PC1, PC2)) + geom_point(aes(col = day, shape = temp)) + theme_bw() + ggtitle("PCA of log2 ADJNSAF values scaled by day*time")
```

Abudance over time of proteins with high PC1 loadings values in the ASCA time-temp interaction effect plot
```{r}
cut_data12_STACKED <- cut_data12
cut_data12_STACKED$day <- gsub("_.*","",rownames(cut_data12))
cut_data12_STACKED$temp <- gsub(".*_","",rownames(cut_data12))

cut_data12_STACKED <- tidyr::gather(cut_data12_STACKED, "protein", "NSAF", 1:44)

ggplot(cut_data12_STACKED, aes(factor(as.integer(day)), NSAF, group = temp)) + geom_point(aes(color = temp)) + geom_line(aes(color = temp)) + facet_wrap(~protein, scale = "free") + ggtitle("Abudance over time of proteins with high PC1 loadings values in the ASCA time-temp interaction effect plot")
```

Heatmap comparison of days x temp norm and non-norm data
```{r}
NSAF_matrix <- NSAF[,-c(1:3)]
rownames(NSAF_matrix) <- paste(NSAF[,2], NSAF[,3], sep = "_")

NSAF_normTxD_matrix <- NSAF_normTxD[,-c(1:2)]
rownames(NSAF_normTxD_matrix) <- paste(NSAF_normTxD[,1], NSAF_normTxD[,2], sep = "_")

library(gtools)
heatmap3(t(NSAF_matrix[mixedsort(rownames(NSAF_matrix)),]), Colv = NA, scale = "row")

heatmap3(t(NSAF_normTxD_matrix[mixedsort(rownames(NSAF_normTxD_matrix)),]), Colv = NA, scale = "row")

```