---
title: "Silo2_3logASCA"
author: "Shelly Trigg"
date: "9/13/2019"
output: html_document
---

```{r}
#exclude day 0 for ASCA
tABACUSdata_avgNSAF_no9D0 <- tABACUSdata_avgNSAF[which(tABACUSdata_avgNSAF$day !=0 & tABACUSdata_avgNSAF$silo !=9),]

#save list of all silo 2 and 3 proteins
all_silo23_prots <- data.frame(rownames(t(tABACUSdata_avgNSAF_no9D0[,grep("CHOYP",colnames(tABACUSdata_avgNSAF_no9D0))])))
write.csv(all_silo23_prots,"~/Documents/GitHub/OysterSeedProject/analysis/ASCA/ASCA_all_proteins_avgADJNSAF/Silo2vs3/Silo2_3_log/all_silo23_prots.csv", row.names = FALSE, quote =FALSE)

```

export data set avg tech. rep. NSAF for all proteins 
```{r,echo=FALSE, eval=FALSE}
```

**Perform ASCA on log2 data**
```{r}
#create matrix to pass to ASCA command, excluding the silo and time info
ASCA_X <- as.matrix(log(tABACUSdata_avgNSAF_no9D0[,-c(1:3)],2))
#create matrix to pass to ASCA command with only the silo and time info
ASCA_F <- as.matrix(tABACUSdata_avgNSAF_no9D0[,c(1:2)])
#perform ASCA
ASCA <- ASCA.Calculate(ASCA_X, ASCA_F, equation.elements = "1,2,12", scaling = FALSE)
```



Here is a summary of the ASCA results (e.g. variance explained by different factors; factor 1= time (days), factor 2 = temperature, interaction = interaction of time and temperature)
```{r}
#print the ASCA summary
ASCA.GetSummary(ASCA)
```


Do a permutations test to see if the factors have a significant effect.
1 = silo
2 = time
3 = silo x time interaction

```{r}
ASCA.permtest <- ASCA.DoPermutationTest(ASCA, perm=1000)
ASCA.permtest
```

### Plot PCAs from ASCA  

**This first plot is the silo effect PCA**
```{r avgNSAF_PCA_timeEffect_plot}
#plot PCA for factor 1, which is silo in this case
ASCA.PlotScoresPerLevel(ASCA, ee = "1", pcs = "1,2")
```

**The next plot is the time effect PCA**
```{r avgNSAF_PCA_timeEffect_plot}
#plot PCA for factor 1, which is time in this case
ASCA.PlotScoresPerLevel(ASCA, ee = "2", pcs = "1,2")
```

**investigate PC1 loadings**
```{r avgNSAF_PCA_tempEffect_PC2loadings, echo = FALSE}
#extract protein names from ASCA data; these will be combined with loadings values; the order is maintained
protnames <- data.frame(colnames(ASCA$data))
#combine protein names with ASCA loadings for PC2 for temperature, since this component showed the greatest separation between temperatures
d <- cbind(protnames, ASCA$`1`$svd$v[,1])
#rename the columns
colnames(d) <- c("protein", "PC1loadings")
#write out Protein x PC1loadings table for GO-mwu
write.csv(d,"~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo23_ASCA_PC1_siloeffect.csv", quote = FALSE, row.names = FALSE)

#make a dataframe of proteins with PC loadings greater than zero for the loadings plot
d_great <- d[which(d$PC1loadings > 0),]
#make a dataframe of proteins with PC loadings less than zero for the loadings plot
d_less <- d[which(d$PC1loadings < 0),]
#write out PC1 loadings for ASCA$2$svd$[,1]
#write.csv(d, "~/Documents/GitHub/paper-OysterSeed-TimeXTemp/Data/Supplementary_Data6.csv", row.names = FALSE, quote = FALSE)
#plot PC1 loadings
plot(d_great[order(d_great$PC1loadings, decreasing = TRUE),2], xlab = "protein", ylab = "PC1 Loadings")
abline(h=0.03, col = "red")
plot(d_less[order(d_less$PC1loadings, decreasing = TRUE),2],xlab = "protein", ylab = "PC1 Loadings")
abline(h=-0.025, col = "red")

```

To pull out proteins affected by temperature based on their influence in seperating treatment groups on PC2 of the temperature PCA, I picked an absolute value loadings threshold of 0.025. This means any protein that had a loadings value > 0.025 or < -0.025 was selected.
```{r, echo = FALSE}
#make list of proteins with temperature PC1 loadings values >= 0.025
cutd <- d[which(d$PC1loadings >= 0.03 | d$PC1loadings <= -0.025),]
#make a list of cutoff proteins with normalized abundance
cut_data <-tABACUSdata_avgNSAF[which(tABACUSdata_avgNSAF$silo !=9),which(colnames(tABACUSdata_avgNSAF) %in% cutd$protein)]
rownames(cut_data) <- paste(tABACUSdata_avgNSAF[which(tABACUSdata_avgNSAF$silo !=9),"silo"],tABACUSdata_avgNSAF[which(tABACUSdata_avgNSAF$silo !=9),"day"], tABACUSdata_avgNSAF[which(tABACUSdata_avgNSAF$silo !=9),"temp"], sep="_")



#######TABLES FOR GO#######
#write out sig proteins
cutd_gmwu <- cbind(data.frame(gsub("\\|","\\.",cutd$protein)), as.numeric(cutd[,2]))

colnames(cutd_gmwu) <- c("protein_ID", "PC1loadings")

cutd_gmwu$PC1loadings <- abs(cutd_gmwu$PC1loadings),10)

write.csv(cutd_gmwu,"~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/ASCA_logNSAFs_FunctionalAnalysis/silo23_sig_ASCA_PC1_siloeffect.csv", quote = FALSE, row.names = FALSE)

cut_data_FC <- data.frame(t(cut_data))

#log transform NSAF values
cut_data_FC <- log2(cut_data_FC)

#calculate silo means
cut_data_FC$silo2_avg <- apply(cut_data_FC[,which(substr(colnames(cut_data_FC),2,2)==2)],1,mean)
cut_data_FC$silo3_avg <- apply(cut_data_FC[,which(substr(colnames(cut_data_FC),2,2)==3)],1,mean)

#calculate p values between silos of log NSAF distributions for for each protein
for (i in 1:nrow(cut_data_FC)){
  test <- wilcox.test(t(cut_data_FC[i,which(substr(colnames(cut_data_FC),2,2)==2)]),t(cut_data_FC[i,which(substr(colnames(cut_data_FC),2,2)==3)]))
  cut_data_FC$p.value[i] <- test$p.value
  cut_data_FC$p.adj[i] <- p.adjust(test$p.value,method = "BH",n=nrow(cut_data_FC))
}



```


**Heatmap of proteins affected by temperature based on temperature effect PC2 loadings value cutoff**
```{r avgNSAF_tempEffectPC1_cutoff0.03&-0.025_heatmap_OrderedByTemp_ShortNames,fig.height=10, fig.width=10, eval = FALSE, echo = FALSE}
#reorder the cut data so that samples are grouped by temperature
cut_data_ord <- cut_data[order(factor(rownames(cut_data),levels = c("0_0_16","2_3_23","2_5_23","2_7_23","2_9_23","2_11_23","2_13_23","3_3_23","3_5_23","3_7_23","3_9_23","3_11_23","3_13_23"))),]

#replace CHOYP names with "entry names"
#transposed to eventually get row names as columns to match entry names to. 
#t.data.frame returns a matrix so I had convert back to data frame for all downstream rearranging because data frames are much easier to work with
cut_data_ord_t <- data.frame(t.data.frame(cut_data_ord))
#make row names which contain the protein IDs a column
cut_data_ord_t <- cbind(rownames(cut_data_ord_t), cut_data_ord_t)
#add the column name
colnames(cut_data_ord_t)[1] <- "Protein.ID"
#convert the Protein ID column to character
cut_data_ord_t[,1] <- as.character(cut_data_ord_t[,1])
#sub pipe for decimal in protein IDs
cut_data_ord_t$Protein.ID <- gsub("\\|","\\.", cut_data_ord_t$Protein.ID)

#add the Entry names
cut_data_ord_t <- merge(data_w_uniprot[,c("Protein.ID","Entry.name.new")],cut_data_ord_t, by = "Protein.ID")
#make the entry names the row names
rownames(cut_data_ord_t) <- cut_data_ord_t[,"Entry.name.new"]
#remove the protein ID and entry name columns since the row names are added
cut_data_ord_t_m <- cut_data_ord_t[,-c(1:2)]
#this step removes the x from the column names while preserving the order of colnames
colnames(cut_data_ord_t_m) <- rownames(cut_data_ord)

#save proteins as a list
silo2_3_sig_temp_prots <- data.frame(rownames(cut_data_ord_t_m), stringsAsFactors = FALSE)
colnames(silo2_3_sig_temp_prots)[1] <- "prots"

#plot the heat map
heatmap3(as.matrix(cut_data_ord_t_m), Colv = NA, cexRow = 0.25)

#make data frames to store heatmap color values
#columns are days/temp; color 23C blue, 29C red, each day a different shade of gray
ColSideColors<-cbind(Silo=c("purple",rep("gold2",6),rep("cyan3",6)), Day=c("white",rep(c("#D9D9D9","#BDBDBD","#969696","#737373","#525252","#252525"),2)))

heatmap3(as.matrix(cut_data_ord_t_m),Colv = NA,cexRow = 0.4, cexCol = 0.8,ColSideColors = ColSideColors,ColAxisColors=1)

```

**extract clusters from heatmap**
```{r}
hm <- as.dist(1-cor(t.data.frame(cut_data_ord_t_m), use="pa"))
hm2 <- hclust(hm, method = 'complete')
plot(hm2)

#Identify clades in heatmap
# define some clusters

#make a list of proteins with cluster IDs based on where the tree is cut
#1.8 gives 2 clusters
mycl <- cutree(hm2, h=1.8)

#assign colors to cluster IDs
clusterCols <- colorRamps::primary.colors(length(unique(mycl)))
myClusterSideBar <- clusterCols[mycl]
heatmap3(as.matrix(cut_data_ord_t_m),cexRow = 0.5, cexCol = 0.5, RowSideColors=myClusterSideBar,ColSideColors = ColSideColors,RowAxisColors=1,ColAxisColors=1,hclustfun=hclust, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), method = "complete", scale = "row", Colv=NA)

#get color name
clusterColor <- lapply(myClusterSideBar, color.id)
clusterColor <- lapply(clusterColor, `[[`,1)
clusterColor <- data.frame(unlist(clusterColor), stringsAsFactors = FALSE)

foo <- cbind(data.frame(mycl), clusterColor)
#add protein column to data frame for merging
foo$Entry.name.new <- rownames(foo)

#add protein column back to NSAF data frame for merging
cut_data_ord_t_m_proteinID <- cut_data_ord_t_m
cut_data_ord_t_m_proteinID$Entry.name.new <- rownames(cut_data_ord_t_m)

clade.prot <- merge(cut_data_ord_t_m_proteinID, foo, by = "Entry.name.new")

colnames(clade.prot)[16] <- "ClusterColor"
colnames(clade.prot)[15] <- "ClusterID"

#remove day 0 from clade prot
clade.prot_noD0 <- clade.prot[,-2]

#normalize within terms (autoscale by row)
#calculate row means and SD
clade.prot_noD0$mean <- apply(clade.prot_noD0[,2:13], 1, mean)
clade.prot_noD0$sd <- apply(clade.prot_noD0[,2:13],1, sd)

#sweep second argument is margin which corresponds to 1 (rows) or 2 (columns); this is for autoscaling by row to make colors relative to one another like the heatmap does
#subtract the row mean from each value
clade.prot_no_D0_norm <- sweep(clade.prot_noD0[,2:13],1,clade.prot_noD0$mean)
#divide each value by the row standard deviation
clade.prot_no_D0_norm <- sweep(clade.prot_no_D0_norm[,1:12],1,clade.prot_noD0$sd, FUN = "/")

clade.prot_no_D0_norm <- cbind(clade.prot_noD0[,c("Entry.name.new", "ClusterID", "ClusterColor")], clade.prot_no_D0_norm)

#save temp sig. proteins with original protein IDs
write.csv(merge(cut_data_ord_t[,1:2], clade.prot_no_D0_norm, by = "Entry.name.new"),"~/Documents/GitHub/OysterSeedProject/analysis/ASCA/ASCA_all_proteins_avgADJNSAF/Silo2vs3/Silo2_3_log/silo2_3_log_tempSig_prots.csv", quote = FALSE, row.names = FALSE)

STACKED_NSAF <- tidyr::gather(clade.prot_no_D0_norm, siloday, NSAF, 4:15)
STACKED_NSAF$siloday <- gsub("_23","",STACKED_NSAF$siloday)
STACKED_NSAF$silo <- gsub("_.*","",STACKED_NSAF$siloday)
STACKED_NSAF$day <- as.factor(gsub(".*_","",STACKED_NSAF$siloday))
str(STACKED_NSAF)
STACKED_NSAF$day <- as.numeric(as.character(STACKED_NSAF$day))

#make column with unique ID
#https://thenode.biologists.com/visualizing-data-one-more-time/education/

STACKED_NSAF$uniq.id <- paste(STACKED_NSAF$Entry.name.new, STACKED_NSAF$silo, sep = "_")

ggplot(STACKED_NSAF,aes(day,NSAF)) + geom_line(aes(group = uniq.id,color = silo),alpha = 0.1)  + stat_smooth(aes(color = silo), method = loess,size = 1) + ylab("mean autoscaled NSAF") + facet_wrap(~ClusterID)  + scale_x_continuous(breaks = c(3,5,7,9,11,13), labels = c(3,5,7,9,11,13)) + scale_color_manual(values=c("gold2", "cyan3")) + theme_bw()

#make mean summary table for plotting summary lines of protein abundances/clade
STACKED_NSAF_mean <- STACKED_NSAF %>%
    group_by(day, silo, ClusterID) %>%
    summarise(n = n(), mean=mean(NSAF), median=median(NSAF), sd=sd(NSAF)) %>%
    mutate(sem = sd / sqrt(n - 1),
      CI_lower = mean + qt((1-0.95)/2, n - 1) * sem,
      CI_upper = mean - qt((1-0.95)/2, n - 1) * sem)

ggplot(STACKED_NSAF_mean, aes(x=day, y=mean, color = silo)) + geom_line(aes(x=day, y=mean, color=silo)) + ylab("mean autoscaled NSAF") + geom_ribbon(aes(ymin=CI_lower,ymax=CI_upper,fill=silo),color="grey70",alpha=0.4)  + facet_wrap(~ClusterID) + scale_color_manual(values=c("gold2", "cyan3")) + scale_fill_manual(values=c("gold2", "cyan3")) + scale_x_continuous(breaks = c(3,5,7,9,11,13), labels = c(3,5,7,9,11,13)) + theme_bw()

```



