---
title: "ASCA_OysterSeedTemp"
author: "Shelly Trigg"
date: "11/5/2018"
output: rmarkdown::github_document
---

Load libraries
```{r}
library(dplyr)
library(tidyr)
library(MetStaT)
library(MetaboAnalystR)
library(ggplot2)
library(heatmap3)
```

read in Oyster Temp. protein data frame
```{r}
data <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/kmeans/Silo3_and_9/silo3and9.csv")
#create silo column
data$silo <- substr(data$Protein,1,1)
#change class of protein column from factor to character
data$Protein <- as.character(unlist(data$Protein))
#remove the silo identifier from the protein names
data$Protein <- substr(data$Protein, 3, nchar(data$Protein))
#Change table orientation so that all abundances are listed in one column 
#and time points are listed down a column
#In this command, we only want to apply it to the timepoint columns that 
#have abundance values and keep the protein and silo columns unchanged
#so we use c(2:9) to select for timepoint columns with abundance values only
data <- tidyr::gather(data, "Time", "abundance", c(2:9))
#Change the table orientation again so that proteins are listed across the top
#This command will condense the proteins so that only unique proteins are listed
#across the top
data <- tidyr::spread(data, "Protein", "abundance")
#change silo column class to numeric
data$silo <- as.numeric(data$silo)
#change time column class to numeric; exclude the X with the substr command
data$Time <- as.numeric(substr(data$Time,2,nchar(data$Time)))

#format and export table for metaboanalyst
data$sample <- paste("S",data$silo,"T",data$Time, sep = "")
#move sample columns to first column
data <- data[,c(ncol(data),1:ncol(data)-1)]
data_choyp <- data[,c(1:3,grep("CHOYP", colnames(data)))]
data_std <- data[,-grep("CHOYP", colnames(data))]
write.csv(data_choyp, "~/Documents/GitHub/OysterSeedProject/analysis/ASCA/ASCA_shellytest/silo3_9_CHOYP_reformat4MetabA.csv", row.names = FALSE, quote = FALSE)
write.csv(data_std, "~/Documents/GitHub/OysterSeedProject/analysis/ASCA/ASCA_shellytest/silo3_9_STD_reformat4MetabA.csv", row.names = FALSE, quote = FALSE)
```


running metaboanalyst in R
```{r}
#read in timeseries data
mSet<-InitDataObjects("conc", "ts", FALSE)
mSet<-SetDesignType(mSet, "time")
mSet<-Read.TextData(mSet, "~/Documents/GitHub/OysterSeedProject/analysis/ASCA/ASCA_shellytest/silo3_9_CHOYP_reformat4MetabA.csv", "rowts", "disc")
mSet<-SanityCheckData(mSet)
#exclude any peptides with missing or zero values
mSet<-RemoveMissingPercent(mSet, percent=0.5)
mSet<-ImputeVar(mSet, method="exclude")
#filter by interquartile range to exclude proteins with near constant values throughout the experiment: 
#mSet<-FilterVariable(mSet, "iqr", "F", 25)

#do mean-centering
mSet<-Normalization(mSet, "NULL", "NULL", "MeanCenter", ratio=FALSE, ratioNum=20)
mSet<-PlotNormSummary(mSet, "norm_CHOYP_nozeros_", "png", 72, width=NA)
#make a data frame out of mean-centered peptides with zero-value peptides excluded
normData <- data.frame(mSet$dataSet$norm)
```

run ASCA with MetStat function
```{r}
#create matrix to pass to ASCA command, excluding the silo and time info
ASCA_normData_X <- as.matrix(normData)
#create matrix to pass to ASCA command with only the silo and time info
ASCA_normData_F <- cbind(as.matrix(as.numeric(substr(rownames(normData),2,2))), as.matrix(as.numeric(substr(rownames(normData),4,length(rownames(normData))))))

#Run ASCA command
ASCA <- ASCA.Calculate(ASCA_normData_X, ASCA_normData_F, equation.elements = "1,2,12", scaling = FALSE)
```

plot loadings for PC1 and PC2 for temperature, time, and their interaction
```{r, echo=FALSE}
ASCA.PlotLoadings(ASCA, ee = "1", pcs="1,2")
ASCA.PlotLoadings(ASCA, ee = "2", pcs="1,2")
ASCA.PlotLoadings(ASCA, ee = "12", pcs="1,2")
```

score plots for PC1 and PC2 for temperature, time, and their interaction
```{r, echo=FALSE}
ASCA.PlotScoresPerLevel(ASCA, ee = "1", pcs = "1,2")
ASCA.PlotScoresPerLevel(ASCA, ee = "2", pcs = "1,2")
ASCA.PlotScoresPerLevel(ASCA, ee = "12", pcs = "1,2")
```

Summary of ASCA
```{r, echo=FALSE}
ASCA.GetSummary(ASCA)
```

Run permutation test (shuffle group labels of samples and see if results are the same)
```{r, echo=FALSE}
ASCA.DoPermutationTest(ASCA, perm=1000)
```

this means that time has a significant effect on molecules, but temperature or time x temp interaction of the factors does not have a significant effect

plot PC1 loadings to find point of diminishing returns for cutoff
print plot PC1 loadings values for proteins with loadings values > 0
```{r, echo=FALSE}
protnames <- data.frame(colnames(ASCA$data))
d <- cbind(protnames, ASCA$`1`$svd$v[,1])
colnames(d) <- c("protein", "PC1loadings")
d_great <- d[which(d$PC1loadings > 0),]
d_less <- d[which(d$PC1loadings < 0),]

plot(d_great[order(d_great$PC1loadings, decreasing = TRUE),2])
plot(d_less[order(d_less$PC1loadings, decreasing = TRUE),2])
```

cutoff loadings at absolute value >= 0.04
make a heatmap to see if proteins cluster based on abundance patterns
```{r}
#make list of proteins with temperature PC1 loadings values >= 0.04
cutd <- d[which(abs(d$PC1loadings) >= 0.04),]

#make a list of cutoff proteins with non-normalized abundance
#first remove pipe
cutDnonNorm <- data_choyp
colnames(cutDnonNorm) <- gsub("\\|","",colnames(cutDnonNorm))
cutDnonNorm <- cutDnonNorm[,c(2:3,which(colnames(cutDnonNorm) %in% cutd$protein))]

#make a list of cutoff proteins with normalized abundance
cutnormD <-normData[,which(colnames(normData) %in% cutd$protein)]
cutnormD$silo <- as.numeric(substr(rownames(cutnormD),2,2))
cutnormD$time <- as.matrix(as.numeric(substr(rownames(cutnormD),4,length(rownames(cutnormD)))))
#reorder columns
cutnormD <- cutnormD[,c(60,61,1:59)]
#order table by silo and time
ord_cutnormD <- cutnormD[order(cutnormD$silo,cutnormD$time),]
#make a heatmap to see if proteins cluster based on abundance over time
heatmap3(as.matrix(t(ord_cutnormD[-c(1,9),3:61])), Colv = NA, cexRow = 0.5)
```

Line plots of mean-centered abundance over time for proteins passing PC1 loadings cutoff
```{r, echo = FALSE, fig.height= 20, fig.width = 10}
#make abundance plots of non-normalized values
cutDnonNorm_long <- tidyr::gather(cutDnonNorm, "protein","abundance", 3:ncol(cutDnonNorm))
cutDnonNorm_long <- merge(cutDnonNorm_long, d)

ggplot(cutDnonNorm_long[which(cutDnonNorm_long$PC1loadings > 0),], aes(x= Time, y = abundance)) + 
  geom_line(aes(colour = as.character(silo))) + facet_wrap(~protein, ncol = 4, scales = "free") +
  theme(strip.text.x = element_text(size = 5)) + ggtitle("all proteins with temperature PC1 loadings values >= 0.04")

ggplot(cutDnonNorm_long[which(cutDnonNorm_long$PC1loadings > 0.1),], aes(x= Time, y = abundance)) + 
  geom_line(aes(colour = as.character(silo))) + facet_wrap(~protein, ncol = 4, scales = "free") +
  theme(strip.text.x = element_text(size = 5)) + ggtitle("all proteins with temperature PC1 loadings values => 0.1")

ggplot(cutDnonNorm_long[which(cutDnonNorm_long$PC1loadings < 0.05),], aes(x= Time, y = abundance)) + 
  geom_line(aes(colour = as.character(silo))) + facet_wrap(~protein, ncol = 4, scales = "free") +
  theme(strip.text.x = element_text(size = 5)) + ggtitle("all proteins with temperature PC1 loadings values =< -0.05")
```


```{r, eval = FALSE, echo = FALSE, fig.height= 20, fig.width = 10}
ord_cutnormD_long <- tidyr::gather(ord_cutnormD, "protein","abundance", 3:ncol(ord_cutnormD))
class(ord_cutnormD_long$silo) <- "character"
ord_cutnormD_long <- merge(ord_cutnormD_long, d)

#make line plots of abundance over time for individual proteins
ggplot(ord_cutnormD_long, aes(x= time, y = abundance)) + geom_line(aes(colour = silo)) + facet_wrap(~protein, ncol = 4) + ggtitle("all proteins with PC1 loadings absolute value >= 0.04") + theme(strip.text.x = element_text(size = 5))
```


```{r, eval = FALSE, echo = FALSE}
ggplot(ord_cutnormD_long[which(ord_cutnormD_long$PC1loadings > 0.1),], aes(x= time, y = abundance)) + geom_line(aes(colour = silo)) + facet_wrap(~protein) + 
ggtitle("all proteins with PC1 loadings > 0.1 (strict)") + theme(strip.text.x = element_text(size = 5))

ggplot(ord_cutnormD_long[which(ord_cutnormD_long$PC1loadings < -0.05),], aes(x= time, y = abundance)) + geom_line(aes(colour = silo)) + facet_wrap(~protein) + ggtitle("all proteins with PC1 loadings < -0.05") +theme(strip.text.x = element_text(size = 5))
```




